---
title: "Does Costco Entry Lower Competitor Fuel Prices in Australia?"
subtitle: "A Synthetic-Control Study for the ACCC"
author: "MGT159 Group Project — Costco Australia"
date: today
format:
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    toc: true
    toc-depth: 2
    number-sections: true
    fig-pos: "H"
    fig-align: center
    fig-width: 7
    fig-height: 4.5
    colorlinks: true
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: "pdf"
    out.width: "100%"
    cache: false
---

```{r setup}
#| cache: false
#| include: false
source("_sc_helpers.R")

library(tidyverse)
library(tidysynth)
library(lubridate)
library(kableExtra)
library(patchwork)
library(scales)
library(glue)

# Reproducibility — the five analysis-ready CSVs live in one public Dropbox
# folder. Appending dl=1 returns the folder contents as a ZIP that we
# download once per session, unzip into tempdir, and read from there.
DROPBOX_FOLDER_URL <- "https://www.dropbox.com/scl/fo/g7lhsenilo4lza6dwnukq/AMIiPcOy1uysFMJtSC3eqrg?rlkey=ozb016wrtnrg9qj9yw789xkyg&st=73xiohe7&dl=1"

cache_dir <- file.path(tempdir(), "costco_sc_inputs")
dir.create(cache_dir, showWarnings = FALSE, recursive = TRUE)

# Headline 5 km / 20 km inputs sit at the root of the Dropbox folder.
# Alt-radius inputs (§4 robustness checks) live under sc_inputs_alt/<variant>/.
required_headline <- c("treated_units.csv", "donor_pool.csv",
                       "treated_metadata.csv", "donor_metadata.csv",
                       "state_median_monthly.csv")
alt_variants      <- c("3km_15km", "8km_30km", "casuarina_10km",
                       "donut_5_20km")
required_alt      <- as.vector(outer(alt_variants,
                                     c("treated_units.csv", "donor_pool.csv",
                                       "treated_metadata.csv"),
                                     function(v, f) file.path("sc_inputs_alt",
                                                              v, f)))
required <- c(required_headline, required_alt)

if (!all(file.exists(file.path(cache_dir, required)))) {
  zip_path <- file.path(cache_dir, "inputs.zip")
  download.file(DROPBOX_FOLDER_URL, zip_path, mode = "wb", quiet = TRUE)
  # junkpaths = FALSE preserves the sc_inputs_alt/<variant>/ subfolder
  # structure for the alt-radius inputs.
  unzip(zip_path, exdir = cache_dir, junkpaths = FALSE)
  missing <- required[!file.exists(file.path(cache_dir, required))]
  if (length(missing) > 0) {
    stop("After unzip, missing files: ", paste(missing, collapse = ", "),
         ". Check the Dropbox folder contents.")
  }
}

treated_units    <- read_csv(file.path(cache_dir, "treated_units.csv"),
                             show_col_types = FALSE)
donor_pool       <- read_csv(file.path(cache_dir, "donor_pool.csv"),
                             show_col_types = FALSE)
treated_metadata <- read_csv(file.path(cache_dir, "treated_metadata.csv"),
                             show_col_types = FALSE)
donor_metadata   <- read_csv(file.path(cache_dir, "donor_metadata.csv"),
                             show_col_types = FALSE)
state_median     <- read_csv(file.path(cache_dir, "state_median_monthly.csv"),
                             show_col_types = FALSE)

# Load the four alt-radius variants. Each variant is a list with three
# tibbles: treated_units, donor_pool, treated_metadata. §4 robustness
# checks index into this list by variant name.
alt_data <- map(alt_variants, function(v) {
  list(
    treated_units    = read_csv(file.path(cache_dir, "sc_inputs_alt", v,
                                          "treated_units.csv"),
                                show_col_types = FALSE),
    donor_pool       = read_csv(file.path(cache_dir, "sc_inputs_alt", v,
                                          "donor_pool.csv"),
                                show_col_types = FALSE),
    treated_metadata = read_csv(file.path(cache_dir, "sc_inputs_alt", v,
                                          "treated_metadata.csv"),
                                show_col_types = FALSE)
  )
}) |> set_names(alt_variants)

# State display order — alphabetical for figures, fixed for tables.
state_levels <- c("NSW", "QLD", "WA")

# Treated-Costco order (south-to-north-ish, matches plan-of-attack figures).
costco_order <- c("Perth Airport", "Casuarina", "Coomera", "Lake Macquarie")
```

\newpage

# Question and motivation {.unnumbered}

We advise the **Australian Competition and Consumer Commission (ACCC)**,
Australia's national competition regulator. The ACCC publishes regular
fuel-price monitoring reports and has flagged ongoing concerns about market
concentration in Australian fuel retailing. One of the ACCC's main tools for
improving competition in concentrated retail markets is reducing entry barriers
for discount entrants — but whether such entrants actually discipline
incumbent pricing is an empirical question, not a foregone conclusion.

We ask:

> **Does the entry of a Costco gas station into an Australian local market
> cause nearby competitor stations to lower their retail fuel prices?**

This is a causal question. We answer it using **synthetic control**, fit
separately for each of the four Australian Costco fuel stations that satisfy
our pre- and post-treatment data-availability rules: Perth Airport (WA, Feb 2020),
Casuarina (WA, Nov 2022), Lake Macquarie (NSW, May 2022), and Coomera (QLD,
May 2023). The complete plan, including identifying assumptions, sample
construction, and the seven pre-committed robustness checks, is in
`deliverables/plan_of_attack_section_{1,2,3}.pdf`. This document is the
*execution* of that plan.

\newpage

# Raw Data Description

## Data sources

We combine three publicly accessible Australian state fuel-price registries —
**NSW FuelCheck** (data.nsw.gov.au), the **Queensland Fuel Price Reporting
Scheme** (data.qld.gov.au), and **WA FuelWatch** (data.wa.gov.au) — with a
hand-collected list of Costco fuel-station opening dates. The three registries
together yield approximately **17.4 million station-day price records across
6,084 unique stations between December 2016 and April 2026**. Each registry
mandates public price reporting by every retail fuel station in the state.

```{r tbl-registries}
#| tbl-cap: "State fuel-price registries used in this analysis."
tibble::tribble(
  ~Registry,        ~State, ~Coverage,                 ~`Raw records`, ~`Unique stations`,
  "NSW FuelCheck",  "NSW",  "Dec 2016 – Jan 2026",     "4,401,971",    "4,813",
  "QLD FPRS",       "QLD",  "Dec 2018 – Dec 2025",     "4,970,650",    "1,883",
  "WA FuelWatch",   "WA",   "Jan 2018 – Apr 2026",     "8,047,153",    "1,423"
) |>
  kbl(booktabs = TRUE, align = "lllrr") |>
  kable_styling(latex_options = c("striped", "hold_position"))
```

NSW FuelCheck and QLD FPRS record price-change events (a row per
station × fuel type × price-update timestamp); WA FuelWatch publishes daily
snapshots (a row per station × fuel type × calendar day). All three are
harmonised in `scripts/synthetic_control_input/build_sc_inputs.py` to a
common station × month panel of mean unleaded price in cents per litre.
Detailed field-level schemas appear in
`deliverables/plan_of_attack_section_1.pdf` §1.1–1.3.

## Treated Costco fuel stations

Ten Costco fuel stations operate (or recently opened) in Australia. Four
satisfy the synthetic-control inclusion rule ($\geq 24$ pre-treatment months and
$\geq 12$ post-treatment months in the relevant state's registry) and form the
treated panel; the remaining six are excluded from formal estimation but
appear here as descriptive context.

```{r tbl-costco-status}
#| tbl-cap: "Australian Costco fuel stations by treatment status."
treated_rows <- treated_metadata |>
  transmute(
    Costco = costco_key,
    State  = state,
    `Treatment date` = format(as.Date(treatment_date), "%b %Y"),
    `Pre months`  = n_pre_months,
    `Post months` = n_post_months,
    Status = "Treated"
  )

# Six excluded Costcos — from plan-of-attack §1.4, hand-collected metadata.
excluded_rows <- tibble::tribble(
  ~Costco,            ~State, ~`Treatment date`, ~`Pre months`, ~`Post months`, ~Status,
  "Marsden Park",     "NSW",  "Jul 2017",        11L,           105L,           "Excluded (pre too short)",
  "Auburn",           "NSW",  "Aug 2025",        107L,            8L,           "Excluded (post too short)",
  "Canberra Airport", "ACT",  "Nov 2023",         0L,            29L,           "Excluded (no pre)",
  "Casula",           "NSW",  "Dec 2016",         4L,           113L,           "Excluded (pre-registry)",
  "North Lakes",      "QLD",  "Jan 2019",         1L,            89L,           "Excluded (pre-registry)",
  "Ipswich",          "QLD",  "Mar 2019",         3L,            87L,           "Excluded (pre-registry)"
)

bind_rows(treated_rows, excluded_rows) |>
  kbl(booktabs = TRUE, align = "llrrrl") |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9)
```

The four treated Costcos give us **staggered treatment timing across three
states** (Feb 2020, May 2022, Nov 2022, May 2023), which is the variation
the synthetic-control design exploits.

## The analysis panel

After ingest, geo-coding (Costco distance), the regular-unleaded filter,
sentinel-price drops, and two-stage station × month → ring/postcode × month
aggregation, our analysis panel contains two pieces:

```{r tbl-panel-summary}
#| tbl-cap: "Analysis-ready panel sizes after sample-construction filters."
panel_summary <- tibble(
  Component = c(
    "Treated 5 km rings × month",
    "Donor postcodes × month",
    "Combined analysis panel"
  ),
  Rows = c(nrow(treated_units),
           nrow(donor_pool),
           nrow(treated_units) + nrow(donor_pool)),
  Units = c(
    glue("{n_distinct(treated_units$costco_key)} treated Costcos"),
    glue("{n_distinct(donor_pool$postcode)} donor postcodes",
         " (NSW = {sum(donor_metadata$state == 'NSW')},",
         " QLD = {sum(donor_metadata$state == 'QLD')},",
         " WA = {sum(donor_metadata$state == 'WA')})"),
    "treated + donor"
  )
)

panel_summary |>
  kbl(booktabs = TRUE,
      format.args = list(big.mark = ",")) |>
  kable_styling(latex_options = c("striped", "hold_position"))
```

The sample-construction funnel (raw rows → aggregated cells → final panel)
is reported in §2.5 as documentary text, since reconstructing it inside this
.qmd would require pulling the full 1.86 GB of raw API archives.

## Distribution of monthly mean unleaded price

```{r fig-price-histogram}
#| fig-cap: "Distribution of monthly mean unleaded price across the combined SC analysis panel (treated 5 km rings + donor postcodes). The distribution is approximately bimodal: the lower mode reflects the 2018–2020 price regime; the higher mode reflects the post-2022 elevated price era following the Russia/Ukraine energy shock."
#| fig-height: 4

panel_combined <- bind_rows(
  treated_units |> transmute(date = as.Date(date), mean_price_cents,
                             segment = "Treated 5 km ring"),
  donor_pool    |> transmute(date = as.Date(date), mean_price_cents,
                             segment = "Donor postcode")
) |>
  filter(!is.na(mean_price_cents))

panel_mean   <- mean(panel_combined$mean_price_cents, na.rm = TRUE)
panel_median <- median(panel_combined$mean_price_cents, na.rm = TRUE)

ggplot(panel_combined, aes(x = mean_price_cents)) +
  geom_histogram(binwidth = 3, fill = "#4C72B0", color = "white", alpha = 0.9) +
  geom_vline(xintercept = panel_mean,   linetype = "dashed",
             color = "#C44E52", linewidth = 0.6) +
  geom_vline(xintercept = panel_median, linetype = "dashed",
             color = "#DD8452", linewidth = 0.6) +
  annotate("label", x = panel_mean,   y = Inf, vjust = 1.5,
           label = glue("mean = {round(panel_mean,   1)} ¢/L"),
           color = "#C44E52", size = 3, label.size = 0) +
  annotate("label", x = panel_median, y = Inf, vjust = 3.5,
           label = glue("median = {round(panel_median, 1)} ¢/L"),
           color = "#DD8452", size = 3, label.size = 0) +
  labs(x = "Postcode-month mean unleaded price (¢/L)",
       y = "Frequency",
       title = glue("Distribution of monthly mean unleaded price (n = {comma(nrow(panel_combined))})")) +
  theme_minimal(base_size = 11)
```

## Monthly median unleaded price by state

```{r fig-state-median-time}
#| fig-cap: "Median of postcode-monthly mean unleaded prices over time, by state. All three states track each other closely: prices fall in 2018–2020 (peaking with the COVID demand collapse in early 2020), rise sharply through 2022 (Russia/Ukraine energy shock), dip during the Mar–Sep 2022 federal fuel-excise cut, and remain elevated through 2026. State-level co-movement supports our use of within-state donor pools to absorb common national shocks."
#| fig-height: 4

state_median |>
  mutate(date = as.Date(date),
         state = factor(state, levels = state_levels)) |>
  ggplot(aes(x = date, y = median_price_cents, color = state)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 0.8, alpha = 0.7) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_color_manual(values = c(NSW = "#4C72B0", QLD = "#DD8452", WA = "#55A868"),
                      name = "State") +
  labs(x = NULL, y = "Median monthly unleaded price (¢/L)",
       title = "Monthly median unleaded price by state, 2017-2026") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "top")
```

## Treated 5 km rings versus state medians

```{r fig-treated-vs-state}
#| fig-cap: "Treated 5 km ring mean unleaded price (colored solid) versus the state median across all reporting postcodes (gray dashed) for each of the four treated Costcos. The vertical red dashed line marks the validated treatment date. Shaded gray: the COVID demand collapse (Mar–Sep 2020); dotted vertical: Russia's invasion of Ukraine (24 Feb 2022); pale-blue shading: the AU federal fuel-excise cut (30 Mar – 28 Sep 2022). The state-median overlay is a descriptive level reference, not the synthetic-control counterfactual."
#| fig-height: 7
#| fig-width: 8

# Shock annotations (Plan of Attack Section 1)
covid_start  <- as.Date("2020-03-01")
covid_end    <- as.Date("2020-09-30")
ukraine      <- as.Date("2022-02-24")
excise_start <- as.Date("2022-03-30")
excise_end   <- as.Date("2022-09-28")

state_colors <- c(NSW = "#C44E52", QLD = "#4C72B0",
                   WA  = "#55A868")

make_panel <- function(costco_name) {
  meta <- treated_metadata |> filter(costco_key == costco_name)
  tr   <- treated_units    |> filter(costco_key == costco_name) |>
    mutate(date = as.Date(date))
  sm   <- state_median     |> filter(state == meta$state) |>
    mutate(date = as.Date(date))

  ggplot() +
    annotate("rect", xmin = covid_start, xmax = covid_end,
             ymin = -Inf, ymax = Inf, fill = "gray80", alpha = 0.5) +
    annotate("rect", xmin = excise_start, xmax = excise_end,
             ymin = -Inf, ymax = Inf, fill = "#B0CFE0", alpha = 0.5) +
    geom_vline(xintercept = ukraine, linetype = "dotted",
               color = "gray30", linewidth = 0.4) +
    geom_line(data = sm, aes(x = date, y = median_price_cents),
              color = "gray50", linetype = "dashed", linewidth = 0.5) +
    geom_line(data = tr, aes(x = date, y = mean_price_cents),
              color = state_colors[[meta$state]], linewidth = 0.7) +
    geom_vline(xintercept = as.Date(meta$treatment_date),
               linetype = "dashed", color = "#C44E52", linewidth = 0.6) +
    scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
    labs(x = NULL, y = "¢/L",
         title = glue("{costco_name} ({meta$state}): 5 km ring vs {meta$state} state median"),
         subtitle = glue("opens {format(as.Date(meta$treatment_date), '%d %b %Y')}")) +
    theme_minimal(base_size = 9) +
    theme(plot.title = element_text(size = 10, face = "bold"),
          plot.subtitle = element_text(size = 8, color = "gray40"))
}

panels <- map(costco_order, make_panel)
wrap_plots(panels, ncol = 1)
```

## Summary statistics

```{r tbl-summary-stats}
#| tbl-cap: "Summary statistics on the SC analysis panel."
price_stats <- panel_combined |>
  summarize(
    n      = n(),
    Mean   = mean(mean_price_cents, na.rm = TRUE),
    SD     = sd(mean_price_cents, na.rm = TRUE),
    Min    = min(mean_price_cents, na.rm = TRUE),
    p25    = quantile(mean_price_cents, 0.25, na.rm = TRUE),
    Median = median(mean_price_cents, na.rm = TRUE),
    p75    = quantile(mean_price_cents, 0.75, na.rm = TRUE),
    Max    = max(mean_price_cents, na.rm = TRUE)
  ) |>
  mutate(Variable = "Monthly mean unleaded price (¢/L)", .before = 1)

dist_stats <- donor_metadata |>
  summarize(
    n      = n(),
    Mean   = mean(min_dist_any_costco_km, na.rm = TRUE),
    SD     = sd(min_dist_any_costco_km, na.rm = TRUE),
    Min    = min(min_dist_any_costco_km, na.rm = TRUE),
    p25    = quantile(min_dist_any_costco_km, 0.25, na.rm = TRUE),
    Median = median(min_dist_any_costco_km, na.rm = TRUE),
    p75    = quantile(min_dist_any_costco_km, 0.75, na.rm = TRUE),
    Max    = max(min_dist_any_costco_km, na.rm = TRUE)
  ) |>
  mutate(Variable = "Donor-postcode distance to nearest Costco (km)", .before = 1)

bind_rows(price_stats, dist_stats) |>
  mutate(across(c(Mean, SD, Min, p25, Median, p75, Max), \(x) round(x, 1))) |>
  kbl(booktabs = TRUE, align = "lrrrrrrrr",
      format.args = list(big.mark = ",")) |>
  kable_styling(latex_options = c("striped", "hold_position", "scale_down"))
```

The mean unleaded price row matches the combined treated + donor sample. The
distance row is computed at the donor-postcode level (one row per postcode);
the plan-of-attack Section 1 reports the analogous statistic at the station
level (n = 5,980, mean $\approx$ 205 km), so the numbers differ in denominator but
agree in shape: most non-Costco stations sit very far from any Costco.

## Data quality and known limitations

- **Price-unit harmonisation.** NSW and WA prices are recorded in cents/L;
  QLD records prices in tenths of cents (e.g. `1899` = 189.9 ¢/L). The build
  pipeline applies the appropriate scaling per registry so the analysis
  panel is uniform in cents/L.
- **Missing coordinates in historical NSW and WA files.** Both registries
  publish historical files without lat/lng; we resolve this with a one-time
  pull from each state's live API, joined to historical records by
  (normalized name, postcode). Stations that closed or rebranded before the
  live-API snapshot fail the join and drop out (52.3% of NSW historical
  stations, 47.3% of WA). The loss falls almost entirely on the donor pool;
  all four treated Costcos join cleanly.
- **NSW Costco rename in May 2022.** Every NSW Costco station was renamed
  in the registry from e.g. `Costco Marsden Park` to
  `Costco Marsden Park (Members only)` on 2 May 2022. The build pipeline
  strips the suffix before matching.
- **Sentinel-price filter.** Quotes outside 50–500 ¢/L are dropped before
  aggregation; this removes obvious out-of-stock placeholders like `9999`
  or `0.01`. After the filter, no postcode-month in the analysis panel
  falls outside 92–259 ¢/L (well inside the plausible Australian range).
- **Casuarina sparse ring.** Casuarina's 5 km ring contains only about
  three contributing stations on average — a documented limitation flagged
  in advance. Robustness Check 3 (Casuarina at 10 km) probes whether this
  noise drives the headline.
- **Perth Airport opens four weeks before COVID.** Its early post-treatment
  months are dominated by pandemic-driven price moves; Robustness Check 4
  (drop Mar–Dec 2020) probes whether this confounds the headline.

\newpage

# Sample Construction

Every choice between raw registry rows and the analysis panel deserves
justification — the headline result is a causal claim about prices, so
opacity at the construction stage would undercut everything downstream.
The subsections below mirror Plan-of-Attack §2(a)–(g) in order. This
section documents the 5 km treated ring / 20 km donor buffer geometry;
alternate geometries appear in §4 Robustness Checks.

## Sample window

The pre-period for each treated Costco runs from the start of its
state's price registry — December 2016 for NSW, January 2018 for WA,
December 2018 for QLD — through the day before that Costco's validated
treatment date (§2.4). The post-period runs from the treatment date
through April 2026, the most recent registry pull. Per-Costco coverage:

```{r tbl-coverage}
#| tbl-cap: "Pre- and post-treatment coverage per treated Costco."
treated_metadata |>
  transmute(
    Costco           = costco_key,
    State            = state,
    `Treatment date` = format(as.Date(treatment_date), "%b %Y"),
    `Pre months`     = n_pre_months,
    `Post months`    = n_post_months
  ) |>
  kbl(booktabs = TRUE, align = "llrrr") |>
  kable_styling(latex_options = c("striped", "hold_position"))
```

Synthetic control fits its weights on the entire pre-period *trajectory*
of the dependent variable, not just on a level. More pre-period
observations tighten identification rather than dilute it — the opposite
of the DID logic in which long pre-periods can introduce structural
breaks into the parallel-trends assumption. Same-state donors experience
the same world-shifts (oil-price regime changes, vehicle fleet turnover,
federal excise adjustments) over the same window, so common shocks are
absorbed into the synthetic by construction. We use as much pre-period
as the data allow.

**COVID.** The 2020 demand collapse and recovery is the only structural
break large enough to warrant explicit handling. Three of the four
treated Costcos opened well after the COVID disruption (Lake Macquarie
May 2022, Casuarina November 2022, Coomera May 2023); for those three
the COVID period sits entirely inside the pre-period and is matched by
the synthetic by construction. **Perth Airport (treated 19 February
2020) opens approximately four weeks before Australia's COVID demand
collapse begins; this is the most fragile case in the analysis.** Its
pre-period is entirely pre-COVID, but its early post-treatment months
are dominated by pandemic-driven price movements rather than competitive
response to Costco's entry. The synthetic counterfactual is built from
WA donor postcodes that experienced the same pandemic shock, so the
post-treatment gap nets out COVID provided the pre-period weights track
the treated trajectory through January 2020. §4 Robustness Check 4
re-estimates Perth Airport with March–December 2020 dropped from both
panels.

## Other sample restrictions

**Treated unit.** For each treated Costco, the treated unit is the
*local market* defined as all non-Costco retail stations within 5 km of
the Costco's coordinates, summarized each month as the simple mean of
those stations' monthly mean unleaded prices. Costco's own price quotes
are excluded from the treated mean — the research question concerns
competitor response, not Costco's own pricing.

**Donor pool.** Each donor unit is a 4-digit Australian postcode,
summarized the same way (monthly mean of non-Costco unleaded price
across all stations within the postcode). Postcodes are the natural
geographic unit of all three state registries and are consistent across
NSW, QLD, and WA. A postcode is excluded from the donor pool if any of
its stations lie within 20 km of any Costco fuel station (treated,
excluded, or otherwise), so donor postcodes are unambiguously outside
any plausible Costco competitive footprint.

**Asymmetric radii (5 km treated, 20 km donor buffer).** The economics
literature on retail fuel-price competition (Houde 2012; Eckert 2013)
finds price effects concentrated within roughly 1–5 miles. 5 km
(~3.1 mi) sits at the conservative end of that range — small enough to
capture genuinely competing stations but not so small as to drop most
of the local market. The 20 km donor buffer is intentionally a 4×
multiple, providing a wide exclusion donut so that no donor postcode is
partly affected by a Costco that isn't the one we're studying. The
asymmetry biases the design *against* finding an effect: a narrow
treated definition shrinks the treated population to the most exposed
stations, and the wide buffer shrinks the donor pool to truly untreated
postcodes.

**Outcome variable.** Regular unleaded — `FuelCode` in
`{U91, E10, P91}` for NSW; `Fuel_Type` in `{Unleaded 91, e10}` for QLD;
`PRODUCT_DESCRIPTION = ULP` for WA. Unleaded is the highest-volume
retail product in Australia, comparable across the three state schemes,
and Costco's headline competitive product.

**Donor-pool measurement-quality filters.** A postcode is additionally
excluded if it averages fewer than 3 unique stations per month (monthly
means based on 1–2 stations are too noisy to serve as credible
counterfactual building blocks) or has fewer than 24 months of any
observed coverage (weights cannot be reliably fit on a short
pre-period).

**Treated-Costco inclusion rule.** A Costco fuel station is treated as
a synthetic-control unit if and only if its registry history contains
both **at least 24 pre-treatment months** *and* **at least 12
post-treatment months** in the relevant state's price registry. Four
of Australia's ten Costco fuel stations satisfy both conditions; six
do not and appear as descriptive context only (§1.2).

```{r tbl-thresholds}
#| tbl-cap: "Threshold-and-defense summary for the SC sample."
tibble::tribble(
  ~Threshold,               ~Value,         ~Justification,                                                                  ~`Robustness check (§4)`,
  "Treated radius",         "at most 5 km",       "Conservative end of IO-literature retail fuel-price spillover range (1–5 mi)",   "Re-run at 3 km and 8 km",
  "Donor buffer",           "> 20 km",      "4× the treated radius; keeps donors fully outside any Costco's competitive zone","Re-run at 15 km and 30 km",
  "Donor stations / month", "at least 3",          "Below 3, monthly mean has high sampling variance",                               "Re-run at at least 2 and at least 5",
  "Donor coverage",         "at least 24 months",  "Minimum pre-period length for stable SC weight fitting",                         "Re-run at at least 36 months",
  "Costco pre-period",      "at least 24 months",  "Same logic, applied to treated",                                                 "Not relaxed (would re-introduce excluded Costcos)",
  "Costco post-period",     "at least 12 months",  "Minimum to estimate a sustained-effect gap",                                     "Not relaxed",
  "Plausible price range",  "50–500 ¢/L",   "Sentinel-value floor and ceiling; AU retail unleaded lives in 80–280 ¢/L",       "None; not an analytic threshold"
) |>
  kbl(booktabs = TRUE, align = "llll") |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9) |>
  column_spec(1, width = "2.8cm") |>
  column_spec(2, width = "2cm") |>
  column_spec(3, width = "5.7cm") |>
  column_spec(4, width = "4.5cm")
```

## Dependent variable integrity

**Construction.** The dependent variable is the monthly mean unleaded
retail price (cents per litre) at the treated 5 km ring or the donor
postcode; see §2.5 for the two-stage construction. @fig-price-histogram shows
the post-aggregation distribution (n = `r comma(nrow(panel_combined))`
postcode-months, range 92–259 ¢/L, approximately bimodal across the
pre-2022 / post-2022 price-regime divide).

**Sentinel-value filter.** Individual price quotes outside 50–500 ¢/L
are dropped before aggregation. This is a data-integrity filter, not
winsorisation: it removes obvious registry artifacts (out-of-stock
placeholder values like `9999` or `0.01`) rather than legitimate tail
observations.

**Missing months.** A postcode-month or treated ring-month with zero
contributing station observations is left as missing (NA), never imputed
as zero, never carried forward. Synthetic-control fitting handles
missing months by ignoring them in the pre-period objective. §1's
data-quality notes report that the largest run of missing months in any
treated panel is three (Lake Macquarie), not material.

**No winsorisation in the main specification.** Two-stage monthly
aggregation smooths individual-quote outliers: an extreme single quote
at one station gets averaged with the rest of that station's month, and
then with the rest of the ring or postcode. If a §4 robustness check
exposes a donor-postcode month driving an estimate, we winsorise the
analysis panel at the 99th percentile and report both specifications
side by side.

## Independent variable integrity

**Treatment indicator.** A binary post-opening flag is constructed per
treated Costco by merging the hand-collected treatment-date table onto
the (Costco × month) panel: 1 from the treatment month forward, 0
before.

**Treatment-date validation.** Registry first-appearance can lag the
true fuel-station opening if a station begins selling fuel before
appearing in the state's reporting feed. Each of the four treated
Costcos' registry-first-appearance dates was cross-checked against
Costco Australia corporate announcements, trade press, and news
coverage. The corroborating source must specifically reference the
*fuel station* rather than the warehouse; Costco Australia warehouses
and fuel stations regularly open weeks apart. If a fuel-station-specific
announced opening date differs from registry first-appearance by more
than 30 days, the announced date supersedes.

```{r tbl-treatment-dates}
#| tbl-cap: "Validated treatment dates: registry first-appearance cross-checked against fuel-station-specific corroboration. Perth Airport is the only revision."
tibble::tribble(
  ~Costco,           ~`Registry first-appearance`, ~`Fuel-station-specific corroboration`,                                                  ~`Delta`,    ~`Treatment date used`,
  "Coomera",         "23 May 2023",                "23 May 2023 (Retail World; explicit fuel-station opening)",                              "0 d",       "23 May 2023 (registry)",
  "Casuarina",       "11 Nov 2022",                "None; trade press only reports the warehouse opening (7 Dec 2022)",                      "n/a",       "11 Nov 2022 (registry)",
  "Perth Airport",   "11 Apr 2020",                "19 Feb 2020 (ACAPMAg: petrol station selling at \\$1.18/L; warehouse opens 19 Mar 2020)", "-52 d",     "\\textbf{19 Feb 2020 (announced)}",
  "Lake Macquarie",  "2 May 2022",                 "None; no fuel-station-specific opening date in public sources",                          "n/a",       "2 May 2022 (registry)"
) |>
  kbl(booktabs = TRUE, align = "lllll", escape = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9) |>
  column_spec(1, width = "2.2cm") |>
  column_spec(2, width = "2.5cm") |>
  column_spec(3, width = "5.5cm") |>
  column_spec(4, width = "1.3cm") |>
  column_spec(5, width = "3cm")
```

Three of the four registry first-appearance dates either match the
fuel-station-specific corroborated date exactly (Coomera) or lack any
fuel-station-specific external date for comparison (Casuarina, Lake
Macquarie); these retain the registry date. **Perth Airport is the only
revision:** the ACAPMAg trade-press report explicitly distinguishes the
petrol station (open 19 Feb 2020, bowser price quoted) from the
warehouse (opening one month later on 19 Mar 2020), placing the
fuel-station opening 52 days before the first FuelWatch record.

**Station coordinates and distance-to-Costco.** Distance is computed as
the great-circle distance from each station's coordinates to the nearest
Costco fuel station's coordinates. NSW and WA historical registry files
do not carry lat/lng, so coordinates are joined from one-time pulls of
the live FuelCheck and FuelWatch APIs by (normalized name, postcode);
QLD's historical files include coordinates inline. Stations that fail
the coordinate join are dropped from analysis — they cannot be
classified into a 5 km ring or a donor postcode without a known
location. Per-registry join failure rates are in §2.6.

## Aggregation

**Why monthly.** Three considerations point at month as the right level
of aggregation. First, NSW and QLD are event-based registries (rows are
price-change events) while WA is a daily-snapshot registry; month is
the lowest common frequency at which all three speak the same language
without having to invent 'daily' observations for the event-based feeds.
Second, the research question is about *sustained* competitive response
to Costco entry, which plays out over weeks-to-months rather than days,
so daily resolution would add noise without identification. Third,
monthly means across multiple stations in a ring or postcode smooth
single-quote outliers without requiring any winsorisation step.

**Two-stage aggregation.** Within each station × month, we compute a
single mean unleaded price across all of that station's price
observations in the month. We then take a simple (unweighted) mean
across stations to produce the (ring × month) treated panel and the
(postcode × month) donor panel. The two-stage structure prevents
stations that change prices more often (and therefore appear more times
in the event-based NSW/QLD feeds) from contributing disproportionately
to the cell mean.

**Carry-forward step (NSW and QLD only).** For NSW and QLD, where rows
are price-change events rather than daily snapshots, we construct each
station's daily price series by carrying its last-observed price forward
to subsequent days within the month until the next observed change.
This is the standard approach for converting price-event data to a
daily series and then to monthly means. WA's daily-snapshot data passes
through unchanged.

## Data merging

Three sources are joined into a single analysis panel:

- **State registries** (NSW FuelCheck, QLD FPRS, WA FuelWatch) provide
  the price observations.
- **Live-API station snapshots** (NSW, WA) and inline coordinates (QLD)
  provide station latitude / longitude, joined to historical registry
  rows by (normalized station name, postcode).
- **Hand-collected Costco metadata** (`data/catalogs/usable_costcos.csv`)
  provides each Costco's lat/lng and validated treatment date, joined
  to treated panels by Costco name.

**Expected losses in the merge.** Two sources of leakage. First,
stations that closed, rebranded, or were re-numbered before the April
2026 live-API pull are absent from that snapshot, so historical rows
that mention them fail the coordinate join. The snapshot misses **2,376
of 4,543 NSW historical stations (52.3%)** and **673 of 1,423 WA
historical stations (47.3%)**; QLD is unaffected because its historical
files carry coordinates inline. The loss falls almost entirely on the
donor pool — all four treated Costcos join cleanly — and
`r n_distinct(donor_pool$postcode)` donor postcodes survive the
downstream geographic and quality filters. Second, station-name typos
and mid-panel renames occasionally break the join; the build pipeline
addresses these programmatically where it can (the May 2022 NSW Costco
"Members only" rename, for example, is stripped before matching).

## Filters and transformations: the funnel from raw to analysis-ready

The funnel below traces the exact row and cell counts from the raw
registry pulls through to the four analysis-ready files in
`data/sc_inputs/`. Row counts in the upper block give way to (station ×
month) cells, then (postcode × month) cells, then postcode counts, then
Costco counts as the panel is aggregated and filtered. **Source:**
[`scripts/synthetic_control_input/build_sc_inputs.py`](../scripts/synthetic_control_input/build_sc_inputs.py).
Bolded rows are the milestone outputs.

```{r tbl-funnel}
#| tbl-cap: "Sample-construction funnel from raw registry rows to the analysis-ready panels in data/sc_inputs/."
tibble::tribble(
  ~Step,                                      ~NSW,         ~QLD,         ~WA,         ~Justification,
  "Raw registry pull (rows)",                 "4,401,971",  "4,970,650",  "8,047,153", "Source data",
  "Filter to regular unleaded (rows)",        "1,982,507",  "997,680",    "2,059,755", "Question is about retail unleaded; premium/diesel are different markets",
  "Drop sentinel prices 50–500 ¢/L (rows)",   "1,982,506",  "995,145",    "2,059,755", "Out-of-stock placeholder removal; nearly nonbinding",
  "Join station coordinates (rows)",          "1,247,728",  "995,145",    "1,289,912", "Required for radius assignment; NSW/WA from live-API snapshot",
  "Aggregate to (station × month) cells",     "156,864",    "70,564",     "68,332",    "Carry-forward for NSW/QLD events; passthrough for WA",
  "Aggregate to (postcode × month) cells",    "42,307",     "18,186",     "18,089",    "Donor candidate panel (561 / 245 / 198 postcodes)",
  "Drop postcodes at most 20 km from any Costco",   "-231 pcds",  "-51 pcds",   "-80 pcds",  "Donors must be outside any Costco's competitive footprint",
  "Drop postcodes < 3 stations / mo avg",     "-226 pcds",  "-138 pcds",  "-82 pcds",  "Monthly mean too noisy below 3 stations",
  "Drop postcodes < 24 months observed",      "-0 pcds",    "-0 pcds",    "-0 pcds",   "Coverage was good; no postcode failed this gate",
  "Final donor pool (postcodes)",             "104",        "56",         "36",        "196 total × ~90 months = 17,578 (postcode × month) cells",
  "Costco eligibility (at least 24 pre, at least 12 post)", "-4 Costcos", "-2 Costcos", "0",         "Inclusion rule applied to all 10 Australian Costco fuel stations",
  "Final treated panel (Costcos)",            "1",          "1",          "2",         "4 Costcos × 5 km ring × ~80–100 months = 375 (Costco × month) cells"
) |>
  kbl(booktabs = TRUE, align = "lllll") |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9) |>
  column_spec(1, width = "5cm") |>
  column_spec(2, width = "1.7cm") |>
  column_spec(3, width = "1.5cm") |>
  column_spec(4, width = "1.5cm") |>
  column_spec(5, width = "5.5cm") |>
  row_spec(c(10, 12), bold = TRUE)
```

**Output.** Four CSVs in `data/sc_inputs/`:
`treated_units.csv` (`r comma(nrow(treated_units))` Costco-month rows),
`donor_pool.csv` (`r comma(nrow(donor_pool))` postcode-month rows),
`treated_metadata.csv` (`r nrow(treated_metadata)` rows), and
`donor_metadata.csv` (`r nrow(donor_metadata)` rows). These four files,
uploaded to the Dropbox folder referenced in the setup chunk, form the
complete input to §3's synthetic-control analysis.

\newpage

# Synthetic-Control Analysis

## Research design

**Question type.** Causal. The identification strategy is **synthetic
control**, fit separately for each of the four treated Costcos.

**Why synthetic control over difference-in-differences.** With four
treated Costcos and 196 candidate donor postcodes, a treated-versus-control
mean comparison is too thin: a single anomalous Costco (e.g. Casuarina,
with only ~3 contributing stations) would dominate the average. Synthetic
control builds a weighted counterfactual for each treated unit
individually — exactly the design the n = 4 case calls for.

**Source of variation.** Costco fuel-station entry into a local market
isn't driven by what's happening to local fuel prices. Site selection
is driven by Costco warehouse logistics, land availability, and Costco's
national rollout schedule — none of which respond to the local
fuel-price trajectory at the chosen site. The variation we exploit is
the staggered timing of the four openings (Feb 2020, May 2022, Nov 2022,
May 2023) across three states.

**Identifying assumptions.** Each is paired with the figure or check
that defends it:

- *Pre-period fit captures unobserved time-varying confounders.*
  Verified visually in @fig-sc-trajectories (pre-period overlap between
  treated and synthetic), quantified in @tbl-fit-quality (pre-period
  tracking error per Costco), and tested by the 12-month holdout in §4
  Robustness Check 5.
- *Donors stay clean.* No donor postcode is partly affected by a Costco
  that isn't the one we're studying. Defended by the 20 km donor buffer
  (4× the treated radius), pre-committed in §2.2, and explicitly tested
  by §4 Robustness Check 7 (spatial placebo on the 5–20 km donut).
- *No anticipation.* Competitor stations do not preemptively cut prices
  ahead of Costco's announced opening. Tested by §4 Robustness Check 1
  (placebo in time, −12 months).
- *Range coverage.* The treated ring's pre-period price level can be
  expressed as a weighted average of donor postcodes. Verified by the
  donor-weight bars in @fig-donor-weights (no negative weights, no
  extreme single-donor concentration).

**What the weights do.** One synthetic-control fit per treated Costco.
For each Costco the donor pool is restricted to **same-state postcodes
only** so the synthetic absorbs state-specific shocks (state excise
rates, regional refining capacity, intra-state shipping costs) by
construction. Weights are chosen to minimize the average gap between
the treated 5 km ring's monthly mean unleaded price and the weighted
donor mean across the pre-period.

**Method commitment.** Standard synthetic-control via R's `tidysynth`
package — the modern tidyverse-friendly successor to the original
`Synth` implementation. **Predictors:** yearly means of the outcome
series over the pre-period (no auxiliary covariates — we have no
postcode-level demographic panel that would justify them).
**Pre-period objective:** minimize the average gap between treated and
synthetic on monthly mean unleaded price (`optimization_window =
pre_period` in `tidysynth::generate_weights`). **Inference:** estimate
uncertainty by treating each donor postcode as a fake treated unit and
measuring the resulting fake effect, summarized as the post/pre
tracking-error ratio and as a 95% CI on the post-treatment gap. All
fitting is driven by [`_sc_helpers.R::fit_one_costco()`](_sc_helpers.R).

## Fitting the four synthetic controls

```{r sc_fits}
#| cache: true
sc_fits <- treated_metadata |>
  arrange(costco_key) |>
  mutate(
    fit = pmap(
      list(costco_key, treatment_date, state),
      function(name, td, st) {
        fit_one_costco(
          treated_units  = treated_units,
          donor_pool     = donor_pool,
          costco_name    = name,
          treatment_date = td,
          state          = st
        )
      }
    )
  )

trajectories <- sc_fits |>
  transmute(
    costco          = costco_key,
    state,
    treatment_date,
    trajectory      = map2(fit, costco_key, trajectory)
  )

per_costco_pointwise_ci <- sc_fits |>
  transmute(
    costco         = costco_key,
    treatment_date,
    ci             = map(fit, permutation_ci_pointwise)
  )
```

The chunk above fits four `tidysynth` models (one per treated Costco),
each with `generate_placebos = TRUE` so the donor-permutation
distribution is computed in the same pass. Caching is enabled
(`cache: true`) so subsequent knits re-use the fits.

## Synthetic-control trajectories per Costco

Four stacked panels of treated vs synthetic trajectories. Solid colored =
actual treated 5 km ring; dashed black = synthetic counterfactual;
vertical dotted red = validated treatment date (§2.4); shaded gray
band = 95% donor-permutation CI on the synthetic, computed by treating
each donor postcode as a placebo treated unit and taking the empirical
2.5–97.5 percentile of donor-side gaps at each time point. The wider
the band, the larger the noise floor at that month — a treated curve
that falls outside the band is evidence the gap is bigger than what
donor-side variability would produce.

```{r fig-sc-trajectories}
#| fig-cap: "Synthetic-control trajectories per treated Costco, stacked for full-width time comparison. Solid colored = treated 5 km ring (actual); dashed black = synthetic counterfactual; vertical dotted red = validated Costco opening; shaded gray = 95% donor-permutation CI on the synthetic."
#| fig-height: 9
#| fig-width: 7

costco_colors <- c(
  "Coomera"        = "#4C72B0",
  "Casuarina"      = "#DD8452",
  "Perth Airport"  = "#55A868",
  "Lake Macquarie" = "#C44E52"
)

make_traj_panel <- function(costco_name) {
  meta <- treated_metadata |> filter(costco_key == costco_name)
  traj <- trajectories |>
    filter(costco == costco_name) |>
    pull(trajectory) |> pluck(1)
  ci <- per_costco_pointwise_ci |>
    filter(costco == costco_name) |>
    pull(ci) |> pluck(1)

  traj_with_ci <- traj |>
    left_join(ci, by = "time_unit") |>
    mutate(
      synth_lo = synth_y + ci_lo,
      synth_hi = synth_y + ci_hi
    )

  ggplot(traj_with_ci, aes(x = time_unit)) +
    geom_ribbon(aes(ymin = synth_lo, ymax = synth_hi),
                fill = "gray80", alpha = 0.5) +
    geom_line(aes(y = synth_y), linetype = "dashed", color = "black",
              linewidth = 0.5) +
    geom_line(aes(y = real_y), color = costco_colors[[costco_name]],
              linewidth = 0.7) +
    geom_vline(xintercept = as.Date(meta$treatment_date),
               linetype = "dotted", color = "#C44E52", linewidth = 0.6) +
    scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
    labs(x = NULL, y = "Mean unleaded price (¢/L)",
         title = glue("{costco_name} ({meta$state})"),
         subtitle = glue("opens {format(as.Date(meta$treatment_date), '%b %Y')}")) +
    theme_minimal(base_size = 9) +
    theme(plot.title    = element_text(size = 10, face = "bold"),
          plot.subtitle = element_text(size = 8,  color = "gray40"))
}

traj_panels <- map(costco_order, make_traj_panel)
wrap_plots(traj_panels, ncol = 1)
```

A reader who sees only @fig-sc-trajectories should be able to read
off the project's conclusion: did competitor prices fall after Costco
entry, and by how much.

## Donor weights per treated Costco

For each Costco we report the postcodes whose synthetic-control weight
exceeds 0.01, sorted in descending order. A counterfactual leaning on
one or two donors is fragile; one drawn from a balanced mix of
geographically sensible donors is credible. Geographic sanity-check:
Coomera should lean on Gold Coast / Brisbane postcodes, Lake Macquarie
on Newcastle / Central Coast, Perth Airport and Casuarina on Perth
metro postcodes.

```{r fig-donor-weights}
#| fig-cap: "Donor-postcode weights per treated Costco (weights > 0.01, sorted descending). A balanced mix across 5-7+ donors with no single donor exceeding ~0.5 is the credibility signal; concentration on 1-2 donors would flag a fragile counterfactual."
#| fig-height: 7
#| fig-width: 8

weights_long <- sc_fits |>
  transmute(
    costco  = costco_key,
    weights = map(fit, donor_weights_top, threshold = 0.01)
  ) |>
  unnest(weights) |>
  mutate(postcode = sub("^pc_", "", unit))

make_weights_panel <- function(costco_name) {
  d <- weights_long |>
    filter(costco == costco_name) |>
    arrange(desc(weight)) |>
    mutate(label = factor(postcode, levels = rev(postcode)))

  if (nrow(d) == 0) {
    return(
      ggplot() + theme_void() +
        labs(title = glue("{costco_name}: no donors with weight > 0.01"))
    )
  }

  ggplot(d, aes(x = weight, y = label)) +
    geom_col(fill = costco_colors[[costco_name]], alpha = 0.85) +
    geom_text(aes(label = sprintf("%.2f", weight)),
              hjust = -0.2, size = 2.8) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.15)),
                       limits = c(0, NA)) +
    labs(x = "Synthetic-control weight", y = "Postcode",
         title = glue("{costco_name}")) +
    theme_minimal(base_size = 9) +
    theme(plot.title = element_text(size = 10, face = "bold"))
}

weight_panels <- map(costco_order, make_weights_panel)
wrap_plots(weight_panels, ncol = 2)
```

## Aggregate event study

The per-Costco picture in @fig-sc-trajectories obscures the aggregate
dynamic. @fig-event-study turns "Costco lowered prices on average" into
"Costco lowered prices by *X* ¢/L within *Y* months and the effect
persisted for *Z*"
— much closer to the answer a regulator can act on. The window is
−24 to +24 relative months, the largest range over which all four
Costcos contribute (limited by Perth Airport's 26 pre-months and
Coomera's 31 post-months).

```{r fig-event-study}
#| fig-cap: "Aggregate event study. X-axis: months relative to Costco opening. Y-axis: mean of (actual - synthetic) gap across the four treated Costcos. Shaded gray: 95% CI from cross-Costco standard error. Vertical dotted red: treatment (month 0)."
#| fig-height: 4.5
#| fig-width: 7

rel_month_gaps <- trajectories |>
  mutate(
    rel = map2(trajectory, treatment_date, relative_month_gaps)
  ) |>
  select(costco, state, rel) |>
  unnest(rel) |>
  filter(rel_month >= -24, rel_month <= 24)

agg_event <- rel_month_gaps |>
  group_by(rel_month) |>
  summarize(
    n        = sum(!is.na(gap)),
    mean_gap = mean(gap, na.rm = TRUE),
    se       = sd(gap, na.rm = TRUE) / sqrt(sum(!is.na(gap))),
    .groups  = "drop"
  ) |>
  mutate(
    ci_lo = mean_gap - 1.96 * se,
    ci_hi = mean_gap + 1.96 * se
  )

ggplot(agg_event, aes(x = rel_month, y = mean_gap)) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi),
              fill = "gray80", alpha = 0.6) +
  geom_hline(yintercept = 0, color = "gray40", linewidth = 0.3) +
  geom_vline(xintercept = 0, linetype = "dotted",
             color = "#C44E52", linewidth = 0.5) +
  geom_line(linewidth = 0.7, color = "#4C72B0") +
  geom_point(size = 1.2, color = "#4C72B0") +
  labs(
    x        = "Months relative to Costco opening",
    y        = "Mean gap (actual - synthetic), ¢/L",
    title    = "Aggregate event study: mean gap by month-since-opening",
    subtitle = "Averaged across the 4 treated Costcos; 95% cross-Costco CI"
  ) +
  theme_minimal(base_size = 11)
```

## Per-Costco effect-size summary

```{r tbl-effect-summary}
#| tbl-cap: "Per-Costco effect-size summary. Mean post-treatment gap = mean of (actual - synthetic) across the post-treatment window. Post/pre MSPE ratio compares post-period mean-squared prediction error to pre-period; values of 4-5+ are the standard threshold for evidence of a real effect."
table_1 <- pmap_dfr(
  list(sc_fits$fit, sc_fits$costco_key, sc_fits$state,
       sc_fits$treatment_date, sc_fits$n_post_months),
  effect_summary
)

table_1 |>
  mutate(ci_str = sprintf("[%.2f, %.2f]", ci_lo, ci_hi)) |>
  transmute(
    Costco             = costco,
    State              = state,
    `Pre mean (¢/L)`   = round(pre_mean_cents, 2),
    `Post gap (¢/L)`   = round(mean_post_gap_cents, 2),
    `\\% red.`         = round(pct_reduction, 2),
    `MSPE ratio`       = round(post_pre_ratio, 2),
    `95\\% CI`         = ci_str,
    `Post mo.`         = post_months
  ) |>
  kbl(booktabs = TRUE, align = "llrrrrrr", escape = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9) |>
  column_spec(1, width = "2.4cm") |>
  column_spec(2, width = "0.8cm") |>
  column_spec(3, width = "1.7cm") |>
  column_spec(4, width = "1.7cm") |>
  column_spec(5, width = "1.2cm") |>
  column_spec(6, width = "1.6cm") |>
  column_spec(7, width = "2.5cm") |>
  column_spec(8, width = "1.2cm")
```

The **post/pre tracking-error ratio** compares how far the synthetic
strays from actual *after* Costco entry to how closely it tracked
*before*. A ratio of 4–5 or higher is the standard threshold for
evidence of a real effect — the post-period gap is several times larger
than the pre-period noise. The **95% CI** is the empirical 2.5–97.5
percentile of post-treatment mean gaps across the donor-permutation
distribution; a CI that does not include zero is evidence against the
null of "no effect."

## Pre-period fit quality

```{r tbl-fit-quality}
#| tbl-cap: "Pre-period fit quality per treated Costco. Lower tracking error means a tighter pre-period match; a tracking error below 1% of the pre-period mean is typically considered a credible fit. Top-3 donor weight share reflects how concentrated the synthetic is on its strongest donors."
table_2 <- pmap_dfr(
  list(sc_fits$fit, sc_fits$costco_key, sc_fits$treatment_date,
       sc_fits$n_pre_months),
  fit_quality_summary
)

table_2 |>
  transmute(
    Costco                              = costco,
    `Pre months`                        = pre_months,
    `Pre tracking error (¢/L)`     = round(pre_rmspe_cents, 3),
    `Tracking error as \\% of pre mean` = round(pre_rmspe_pct, 2),
    `Top-3 donor weight share`          = round(top3_donor_share, 2)
  ) |>
  kbl(booktabs = TRUE, align = "lrrrr", escape = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9) |>
  column_spec(3, width = "3cm") |>
  column_spec(4, width = "3.5cm") |>
  column_spec(5, width = "3cm")
```

Casuarina's pre-period tracking error is expected to be visibly larger
than the others because the 5 km Casuarina ring contains only ~3
contributing stations on average (§1's documented limitation).
Surfacing this transparently here is part of presenting the limitation
rather than hiding it.

\newpage

# Robustness Checks

The biggest sources of fragility in the headline analysis are (i) the
5 km / 20 km geographic geometry, (ii) the COVID period coinciding with
Perth Airport's opening, (iii) the noise in Casuarina's sparse 5 km
ring, (iv) the standard synthetic-control concern that pre-period fit
can be gamed by over-flexible donor selection, and (v) the lack of
formal uncertainty quantification on a four-Costco point estimate. The
seven pre-committed checks below directly probe these concerns. Checks
are numbered to match plan-of-attack §3(c). Checks 2, 3, and 7 use
alternate-radius SC inputs built by
[`scripts/synthetic_control_input/build_sc_inputs_alt_radii.py`](../scripts/synthetic_control_input/build_sc_inputs_alt_radii.py)
and pulled from the same Dropbox folder as the headline inputs.

## Placebo in time (RC1)

**Concern.** No anticipation: competitor stations do not pre-emptively
cut prices ahead of Costco's announced opening. If a 'ban-style effect'
appears in the placebo window before Costco actually opens, the
post-treatment gap reflects pre-existing trends or anticipation, not
Costco entry.

**Procedure.** For each treated Costco, backdate treatment by 12 months
to define a placebo opening date. Refit synthetic control on the
truncated panel — all months strictly before the real treatment date —
so no real post-treatment observations contaminate the placebo window.
Compute the mean placebo gap over the 12 months from the placebo date
to the real treatment date, and report the donor-permutation 95% CI.

```{r rc1_placebo}
#| cache: true
# The placebo fit truncates the pre-period by 12 months, which often
# makes per-year predictors collinear (kernlab::ipop singular-matrix
# error). fit_one_costco_robust handles the fallback automatically.
rc1_fits <- treated_metadata |>
  arrange(costco_key) |>
  mutate(
    placebo_date = as.Date(treatment_date) %m-% months(12),
    fit = pmap(
      list(costco_key, treatment_date, placebo_date, state),
      function(name, td, pd, st) {
        # Truncate the treated panel to dates strictly before the REAL
        # treatment date. fit_one_costco's internal donor filter
        # restricts donors to the same grid, so a separate donor
        # truncation isn't needed.
        tu_pre <- treated_units |>
          filter(as.Date(date) < as.Date(td))
        fit_one_costco_robust(
          treated_units  = tu_pre,
          donor_pool     = donor_pool,
          costco_name    = name,
          treatment_date = pd,
          state          = st
        )
      }
    )
  )
```

```{r tbl-rc1}
#| tbl-cap: "RC1 placebo-in-time: mean (actual - synthetic) gap over the 12 months between the placebo date and the real treatment date. A gap near zero — and a 95% CI that includes zero — is the pass criterion."
rc1_table <- pmap_dfr(
  list(rc1_fits$fit, rc1_fits$costco_key, rc1_fits$state,
       rc1_fits$treatment_date, rc1_fits$placebo_date),
  function(fit, name, st, real_td, placebo_td) {
    if (is.null(fit)) {
      return(tibble(
        Costco              = name,
        State               = st,
        `Placebo date`      = format(as.Date(placebo_td), "%b %Y"),
        `Real date`         = format(as.Date(real_td),    "%b %Y"),
        `Mean placebo gap (¢/L)` = NA_real_,
        `95\\% CI`          = "(fit failed)"
      ))
    }
    traj <- trajectory(fit, name)
    pd_anchor <- month_anchor(placebo_td)
    td_anchor <- month_anchor(real_td)
    placebo_window <- traj |>
      filter(.data$time_unit >= pd_anchor,
             .data$time_unit <  td_anchor)
    placebo_gap <- mean(placebo_window$gap, na.rm = TRUE)
    ci <- permutation_ci_mean_gap(fit, placebo_td)
    tibble(
      Costco              = name,
      State               = st,
      `Placebo date`      = format(as.Date(placebo_td), "%b %Y"),
      `Real date`         = format(as.Date(real_td),    "%b %Y"),
      `Mean placebo gap (¢/L)` = round(placebo_gap, 2),
      `95\\% CI`          = sprintf("[%.2f, %.2f]", ci[["lo"]], ci[["hi"]])
    )
  }
)

rc1_table |>
  kbl(booktabs = TRUE, align = "llllrl", escape = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9)
```

**Interpretation.** The pass criterion is a placebo gap near zero with
a 95% CI that includes zero — meaning donor-side variability is wider
than the placebo gap, so the headline post-treatment gap cannot be
attributed to a pre-existing downward trend or to anticipation in the
12 months before Costco's actual opening.

## Alternative radii (RC2)

**Concern.** The 5 km treated radius and 20 km donor buffer are
defensible (Houde 2012, Eckert 2013 place price effects within 1–5
miles; the 4× buffer keeps donors fully outside the competitive zone)
but discretionary. An effect that exists at 5 km but vanishes at 3 km
or 8 km would be fragile. This is the most important sensitivity test
because the radius geometry *is* the design.

**Procedure.** Re-build SC inputs at (3 km treated, 15 km donor buffer)
and (8 km treated, 30 km donor buffer). Refit each Costco at each
geometry and report per-Costco gaps in a sensitivity table.

```{r rc2_alt_radii}
#| cache: true
# Helper: refit all four Costcos using one alt-radius variant's panels.
fit_variant <- function(variant_name) {
  v <- alt_data[[variant_name]]
  v$treated_metadata |>
    arrange(costco_key) |>
    mutate(
      fit = pmap(
        list(costco_key, treatment_date, state),
        function(name, td, st) {
          fit_one_costco_robust(
            treated_units  = v$treated_units,
            donor_pool     = v$donor_pool,
            costco_name    = name,
            treatment_date = td,
            state          = st
          )
        }
      )
    )
}

rc2_3km_15km <- fit_variant("3km_15km")
rc2_8km_30km <- fit_variant("8km_30km")

summarise_variant <- function(fit_tbl, label) {
  fit_tbl_ok <- fit_tbl |> filter(!map_lgl(fit, is.null))
  if (nrow(fit_tbl_ok) == 0) {
    return(tibble(spec = label, costco = NA_character_,
                  pre_mean_cents = NA_real_, mean_post_gap_cents = NA_real_,
                  pct_reduction = NA_real_, post_pre_ratio = NA_real_,
                  ci_lo = NA_real_, ci_hi = NA_real_, post_months = NA_integer_))
  }
  pmap_dfr(
    list(fit_tbl_ok$fit, fit_tbl_ok$costco_key, fit_tbl_ok$state,
         fit_tbl_ok$treatment_date, fit_tbl_ok$n_post_months),
    effect_summary
  ) |>
    mutate(spec = label, .before = 1)
}

rc2_combined <- bind_rows(
  table_1                                        |> mutate(spec = "5 km / 20 km (headline)"),
  summarise_variant(rc2_3km_15km, "3 km / 15 km"),
  summarise_variant(rc2_8km_30km, "8 km / 30 km")
)
```

```{r tbl-rc2}
#| tbl-cap: "RC2 sensitivity to radius geometry. Per-Costco gaps at the headline 5/20 km specification and at the two alternate geometries (3/15 km, 8/30 km). The headline conclusion is robust if sign and approximate magnitude survive at both alt radii."
rc2_combined |>
  transmute(
    Specification     = spec,
    Costco            = costco,
    `Pre mean (¢/L)` = round(pre_mean_cents, 2),
    `Post gap (¢/L)` = round(mean_post_gap_cents, 2),
    `\\% red.`       = round(pct_reduction, 2),
    `MSPE ratio`     = round(post_pre_ratio, 2),
    `95\\% CI`       = sprintf("[%.2f, %.2f]", ci_lo, ci_hi)
  ) |>
  kbl(booktabs = TRUE, align = "llrrrrl", escape = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 8) |>
  collapse_rows(columns = 1, valign = "top")
```

**Interpretation.** Treat the 5/20 km row as the reference. For each
Costco we check (a) that the post-treatment gap stays negative at both
alt radii and (b) that magnitudes do not collapse. A sign flip at any
alt radius — or a magnitude reduction of more than ~50% — flags that
Costco's headline estimate as geometry-sensitive.

## Casuarina at 10 km treated radius (RC3)

**Concern.** Casuarina's 5 km ring contains only ~3 contributing
stations on average (documented in §1's data-quality notes). A wider ring captures more
signal at the cost of including some less-exposed stations. If the
headline Casuarina estimate flips or shrinks substantially at 10 km,
the per-Costco estimate may be driven by individual-station noise
rather than competitive response.

**Procedure.** Re-fit Casuarina specifically at a 10 km treated radius
(donor buffer remains 20 km). Compare gap and pre-period tracking
error to the 5 km baseline.

```{r rc3_casuarina_10km}
#| cache: true
v3 <- alt_data[["casuarina_10km"]]
cas_meta <- v3$treated_metadata |> filter(costco_key == "Casuarina")

rc3_fit <- fit_one_costco_robust(
  treated_units  = v3$treated_units,
  donor_pool     = v3$donor_pool,
  costco_name    = "Casuarina",
  treatment_date = cas_meta$treatment_date,
  state          = "WA"
)

rc3_effect  <- if (is.null(rc3_fit)) NULL else effect_summary(
  rc3_fit, "Casuarina", "WA",
  cas_meta$treatment_date, cas_meta$n_post_months
)
rc3_quality <- if (is.null(rc3_fit)) NULL else fit_quality_summary(
  rc3_fit, "Casuarina",
  cas_meta$treatment_date, cas_meta$n_pre_months
)
```

```{r tbl-rc3}
#| tbl-cap: "RC3 Casuarina at the headline 5 km radius vs the 10 km robustness radius. Pre-period tracking error is expected to fall under the wider ring (more contributing stations smooth idiosyncratic noise); the question is whether the post-treatment gap moves correspondingly."
casuarina_headline_effect  <- table_1 |> filter(costco == "Casuarina")
casuarina_headline_quality <- table_2 |> filter(costco == "Casuarina")

rc3_row <- if (is.null(rc3_effect)) {
  tibble(
    Spec               = "10 km (RC3)",
    `Pre mean (¢/L)`  = NA_real_,
    `Post gap (¢/L)`  = NA_real_,
    `\\% red.`        = NA_real_,
    `MSPE ratio`      = NA_real_,
    `Pre tracking err (¢/L)` = NA_real_,
    `95\\% CI`        = "(fit failed)"
  )
} else {
  tibble(
    Spec               = "10 km (RC3)",
    `Pre mean (¢/L)`  = round(rc3_effect$pre_mean_cents, 2),
    `Post gap (¢/L)`  = round(rc3_effect$mean_post_gap_cents, 2),
    `\\% red.`        = round(rc3_effect$pct_reduction, 2),
    `MSPE ratio`      = round(rc3_effect$post_pre_ratio, 2),
    `Pre tracking err (¢/L)` = round(rc3_quality$pre_rmspe_cents, 2),
    `95\\% CI`        = sprintf("[%.2f, %.2f]",
                                rc3_effect$ci_lo, rc3_effect$ci_hi)
  )
}

bind_rows(
  tibble(
    Spec               = "5 km (headline)",
    `Pre mean (¢/L)`  = round(casuarina_headline_effect$pre_mean_cents, 2),
    `Post gap (¢/L)`  = round(casuarina_headline_effect$mean_post_gap_cents, 2),
    `\\% red.`        = round(casuarina_headline_effect$pct_reduction, 2),
    `MSPE ratio`      = round(casuarina_headline_effect$post_pre_ratio, 2),
    `Pre tracking err (¢/L)` = round(casuarina_headline_quality$pre_rmspe_cents, 2),
    `95\\% CI`        = sprintf("[%.2f, %.2f]",
                                casuarina_headline_effect$ci_lo,
                                casuarina_headline_effect$ci_hi)
  ),
  rc3_row
) |>
  kbl(booktabs = TRUE, align = "lrrrrrl", escape = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 8)
```

**Interpretation.** Pre-period tracking error should drop at 10 km
because the wider ring includes more contributing stations and the
monthly mean is less noisy. If the post-treatment gap is similar at
both radii, the headline Casuarina estimate is not driven by
individual-station noise. A large swing — especially a sign flip —
would mean the 5 km point estimate is unreliable and should be
de-emphasized in the headline narrative.

## Perth Airport with the COVID period removed (RC4)

**Concern.** Perth Airport opens 19 February 2020, approximately four
weeks before Australia's COVID demand collapse. Early post-treatment
months are dominated by pandemic-driven price movements rather than
competitive response. COVID-era compression of fuel prices may suppress
the actual - synthetic gap; the COVID-removed estimate may be larger
in magnitude than the full-window gap.

**Procedure.** Re-fit Perth Airport with March–December 2020 dropped
from both the treated and donor panels. Compare the resulting gap to
the headline Perth Airport gap.

```{r rc4_perth_covid}
#| cache: true
covid_drop_start <- as.Date("2020-03-01")
covid_drop_end   <- as.Date("2020-12-31")

is_covid_dropped <- function(d) {
  d <- as.Date(d)
  d >= covid_drop_start & d <= covid_drop_end
}

treated_pa_no_covid <- treated_units |>
  filter(!(costco_key == "Perth Airport" & is_covid_dropped(date)))

donor_no_covid <- donor_pool |>
  filter(!is_covid_dropped(date))

pa_meta <- treated_metadata |> filter(costco_key == "Perth Airport")

rc4_fit <- fit_one_costco_robust(
  treated_units  = treated_pa_no_covid,
  donor_pool     = donor_no_covid,
  costco_name    = "Perth Airport",
  treatment_date = pa_meta$treatment_date,
  state          = "WA"
)
```

```{r fig-rc4}
#| fig-cap: "Perth Airport synthetic-control trajectory with March-December 2020 dropped from both treated and donor panels. Solid green = treated 5 km ring; dashed black = synthetic counterfactual; vertical dotted red = validated treatment date (19 Feb 2020). The gap in early 2020 is the dropped window."
#| fig-height: 4
#| fig-width: 7

if (is.null(rc4_fit)) {
  ggplot() + theme_void() +
    labs(title = "RC4 Perth Airport fit failed",
         subtitle = "See the warning emitted in the rc4_perth_covid chunk for details.")
} else {
  rc4_traj <- trajectory(rc4_fit, "Perth Airport") |>
    filter(!is_covid_dropped(time_unit))

  ggplot(rc4_traj, aes(x = time_unit)) +
    geom_line(aes(y = synth_y), linetype = "dashed", color = "black",
              linewidth = 0.5) +
    geom_line(aes(y = real_y), color = "#55A868", linewidth = 0.7) +
    geom_vline(xintercept = as.Date(pa_meta$treatment_date),
               linetype = "dotted", color = "#C44E52", linewidth = 0.6) +
    annotate("rect", xmin = covid_drop_start, xmax = covid_drop_end,
             ymin = -Inf, ymax = Inf, fill = "gray80", alpha = 0.5) +
    annotate("label", x = covid_drop_start + (covid_drop_end - covid_drop_start) / 2,
             y = Inf, vjust = 1.3, label = "Mar-Dec 2020 dropped",
             color = "gray30", size = 2.8, label.size = 0) +
    scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
    labs(x = NULL, y = "Mean unleaded price (¢/L)",
         title = "Perth Airport: SC fit with COVID period removed",
         subtitle = "Treatment: 19 Feb 2020; dropped window: Mar-Dec 2020") +
    theme_minimal(base_size = 10)
}
```

```{r tbl-rc4}
#| tbl-cap: "RC4 Perth Airport with the COVID period removed, compared to the headline full-window estimate. A larger gap in the COVID-removed specification indicates that pandemic compression of fuel prices suppressed the headline estimate."
pa_headline  <- table_1 |> filter(costco == "Perth Airport")

rc4_summary <- if (is.null(rc4_fit)) {
  pa_headline |>
    mutate(
      mean_post_gap_cents = NA_real_, pct_reduction = NA_real_,
      post_pre_ratio = NA_real_, ci_lo = NA_real_, ci_hi = NA_real_
    )
} else {
  effect_summary(rc4_fit, "Perth Airport", "WA",
                 pa_meta$treatment_date,
                 pa_meta$n_post_months - 10)
}

bind_rows(
  pa_headline  |> mutate(spec = "Headline (full window)"),
  rc4_summary  |> mutate(spec = "RC4 (Mar-Dec 2020 dropped)")
) |>
  transmute(
    Specification        = spec,
    `Pre mean (¢/L)`    = round(pre_mean_cents, 2),
    `Post gap (¢/L)`    = round(mean_post_gap_cents, 2),
    `\\% red.`          = round(pct_reduction, 2),
    `MSPE ratio`        = round(post_pre_ratio, 2),
    `95\\% CI`          = sprintf("[%.2f, %.2f]", ci_lo, ci_hi),
    `Post mo.`          = post_months
  ) |>
  kbl(booktabs = TRUE, align = "lrrrrrr", escape = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9)
```

**Interpretation.** If the COVID-removed gap is *more negative* than
the headline, COVID-era price compression suppressed the actual -
synthetic divergence and the headline understates the true effect at
Perth Airport. If the two estimates are close, the headline already
nets out the pandemic by leaning on within-state donors that
experienced the same shock.

## LOOCV donor selection with 12-month holdout (RC5)

**Concern.** Pre-period fit can be gamed by over-flexible donor
selection — the synthetic 'memorises' the training series without
generalising. The honest test is an out-of-sample holdout: did the
weights learn structural similarity, or did they memorise noise?

**Procedure.** For each Costco, refit the synthetic on the pre-period
*minus the last 12 months* (the training window). The untouched
12-month holdout sits between the training window and the real
treatment date. Without looking at any post-treatment data, score
whether the synthetic continues to track the treated series across the
holdout: holdout RMSPE comparable to training RMSPE means the fit
generalises; a much larger holdout RMSPE means the fit was memorising.

The plan-of-attack also describes a leave-one-out cross-validation over
donor-pool sizes N ∈ {5, 10, 20, 50, full} to choose how many donors to
retain. tidysynth has no native N-restriction primitive, and writing
one introduces an additional layer of researcher discretion the
holdout test is itself trying to remove — so we report the holdout
diagnostic directly on the full donor pool and document the
simplification here.

```{r rc5_holdout}
#| cache: true
# The training window is the pre-period minus its last 12 months. For
# Perth Airport (only 26 pre-months in the headline) the training
# window is ~14 months, too short to support per-year predictors —
# the single overall pre-period mean predictor is the robust choice
# for all four Costcos here.
rc5_fits <- treated_metadata |>
  arrange(costco_key) |>
  mutate(
    train_end    = as.Date(treatment_date) %m-% months(12),
    fit          = pmap(
      list(costco_key, treatment_date, state, train_end),
      function(name, td, st, te) {
        fit_one_costco_robust(
          treated_units  = treated_units,
          donor_pool     = donor_pool,
          costco_name    = name,
          treatment_date = td,
          state          = st,
          fit_end_date   = te
        )
      }
    )
  )
```

```{r tbl-rc5}
#| tbl-cap: "RC5 12-month holdout. Training window = pre-period minus last 12 months. Holdout = untouched last 12 months of pre-period. Post = post-treatment window. A holdout-to-training RMSPE ratio near 1 means the synthetic generalises out-of-sample; a ratio much larger than 1 means the headline pre-period fit was memorising training noise."
rc5_table <- pmap_dfr(
  list(rc5_fits$fit, rc5_fits$costco_key, rc5_fits$treatment_date,
       rc5_fits$train_end),
  function(fit, name, td, te) {
    if (is.null(fit)) {
      return(tibble(
        Costco                 = name,
        `Train months`         = NA_integer_,
        `Train RMSPE (¢/L)`    = NA_real_,
        `Holdout RMSPE (¢/L)`  = NA_real_,
        `Post RMSPE (¢/L)`     = NA_real_,
        `Holdout / Train`      = NA_real_
      ))
    }
    traj <- trajectory(fit, name)
    train_end_anchor <- month_anchor(te)
    treat_anchor     <- month_anchor(td)

    train_traj   <- traj |> filter(.data$time_unit <  train_end_anchor)
    holdout_traj <- traj |> filter(.data$time_unit >= train_end_anchor,
                                   .data$time_unit <  treat_anchor)
    post_traj    <- traj |> filter(.data$time_unit >= treat_anchor)

    rmspe <- function(d) sqrt(mean((d$real_y - d$synth_y)^2, na.rm = TRUE))
    tibble(
      Costco                 = name,
      `Train months`         = nrow(train_traj),
      `Train RMSPE (¢/L)`    = round(rmspe(train_traj), 2),
      `Holdout RMSPE (¢/L)`  = round(rmspe(holdout_traj), 2),
      `Post RMSPE (¢/L)`     = round(rmspe(post_traj), 2),
      `Holdout / Train`      = round(rmspe(holdout_traj) / rmspe(train_traj), 2)
    )
  }
)

rc5_table |>
  kbl(booktabs = TRUE, align = "lrrrrr", escape = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9)
```

**Interpretation.** The holdout-to-training ratio is the headline
diagnostic. Values near 1 mean the synthetic learned a structural match;
values substantially > 1 (say, > 3-4) mean the pre-period fit reported
in @tbl-fit-quality may have been generous about generalisation. The
post-RMSPE column is shown for orientation: it should be much larger
than holdout-RMSPE if the post-treatment gap is a genuine treatment
effect rather than fit drift.

## 95% CIs via donor-permutation inference (RC6)

**Concern.** Point estimates without uncertainty bands can make a
modest effect look airtight. Casuarina's documented noise problem
(@tbl-fit-quality) is the case most at risk of overclaiming precision.

**Procedure.** Treat each donor postcode as a placebo treated unit,
refit synthetic control on the remaining donors, and compute the
resulting placebo gap. Across all donor placebos, take the empirical
2.5–97.5 percentile of post-treatment gaps as a 95% CI on the real
treatment effect.

**Implementation.** This is the same donor-permutation distribution
that powers the shaded bands in @fig-sc-trajectories and the `95% CI`
column in @tbl-effect-summary. Both were computed as part of the §3
fits (`generate_placebos = TRUE` in [_sc_helpers.R](_sc_helpers.R)).
A CI that excludes zero is evidence against the no-effect null;
Casuarina-specific results will be read transparently in either
direction. No new computation is required for this check beyond what
already powers Figure 1 and Table 1.

## Spatial placebo: 5–20 km donut as faux donors (RC7)

**Concern.** The 20 km donor buffer is meant to keep donors fully
outside any Costco's competitive zone, but this assumption is never
verified in the headline analysis. The 5–20 km donut — stations
currently excluded from both treated and donor groups — is a natural
laboratory: treat the donut as a faux treated unit and check whether
synthetic-control spuriously finds a 'Costco effect' there.

Two outcomes are both informative. (a) The donut shows an effect:
Costco's competitive footprint extends beyond 5 km and the headline
estimate is conservative (the true effect is probably larger).
(b) The donut shows no effect: the Costco impact is concentrated
within 5 km and the 20 km buffer is more than sufficient.

```{r rc7_donut}
#| cache: true
vd <- alt_data[["donut_5_20km"]]

rc7_fits <- vd$treated_metadata |>
  arrange(costco_key) |>
  mutate(
    fit = pmap(
      list(costco_key, treatment_date, state),
      function(name, td, st) {
        fit_one_costco_robust(
          treated_units  = vd$treated_units,
          donor_pool     = vd$donor_pool,
          costco_name    = name,
          treatment_date = td,
          state          = st
        )
      }
    )
  )
```

```{r tbl-rc7}
#| tbl-cap: "RC7 spatial placebo. The headline 5 km ring and the 5-20 km donut share the same >20 km donor pool. If the donut shows a Costco-style negative gap, the headline 5 km estimate is conservative (the effect extends beyond 5 km); if the donut gap is near zero with a CI that includes zero, the headline is contained within the 5 km footprint and the 20 km buffer is safely conservative."
rc7_fits_ok <- rc7_fits |> filter(!map_lgl(fit, is.null))
rc7_effects <- if (nrow(rc7_fits_ok) == 0) {
  tibble(costco = character(), state = character(),
         pre_mean_cents = double(), mean_post_gap_cents = double(),
         pct_reduction = double(), post_pre_ratio = double(),
         ci_lo = double(), ci_hi = double(), post_months = integer())
} else {
  pmap_dfr(
    list(rc7_fits_ok$fit, rc7_fits_ok$costco_key, rc7_fits_ok$state,
         rc7_fits_ok$treatment_date, rc7_fits_ok$n_post_months),
    effect_summary
  )
}

bind_rows(
  table_1     |> mutate(spec = "5 km ring (headline)"),
  rc7_effects |> mutate(spec = "5-20 km donut (RC7)")
) |>
  transmute(
    Specification     = spec,
    Costco            = costco,
    `Pre mean (¢/L)` = round(pre_mean_cents, 2),
    `Post gap (¢/L)` = round(mean_post_gap_cents, 2),
    `\\% red.`       = round(pct_reduction, 2),
    `MSPE ratio`     = round(post_pre_ratio, 2),
    `95\\% CI`       = sprintf("[%.2f, %.2f]", ci_lo, ci_hi)
  ) |>
  kbl(booktabs = TRUE, align = "llrrrrl", escape = FALSE) |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 8) |>
  collapse_rows(columns = 1, valign = "top")
```

**Interpretation.** The donut's post-treatment gap is the diagnostic.
A statistically clean zero — gap small, CI straddling zero, MSPE ratio
near 1 — means the headline 5 km treated definition captures the
Costco footprint without leakage. A clearly negative donut gap means
the footprint extends past 5 km and the headline is conservative
(true effect is probably larger). A clearly positive donut gap would
suggest substitution toward donut stations, which would be unusual
but worth flagging.

## Summary

The headline conclusion in §3.5 should be read jointly with this
section. RC1 checks for anticipation; RC2 / RC3 probe the radius
geometry; RC4 isolates Perth Airport's pandemic confound; RC5 is the
out-of-sample holdout test; RC6 attaches uncertainty bands (already
reported in @tbl-effect-summary and @fig-sc-trajectories); RC7 spatially
places the headline within / beyond the 5 km treated definition. A
Costco-level estimate weakens whenever a check inverts the sign of its
gap or pushes the post/pre tracking-error ratio below roughly 3; the
corresponding row of @tbl-effect-summary should then be discounted in
the §5 recommendation.

\newpage

# Recommendation

## What the data actually shows

```{r recommendation_setup}
#| include: false
# Pull headline numbers from §3 so the prose stays consistent with
# whatever the fits produce.
headline_mean_gap_cents <- mean(table_1$mean_post_gap_cents, na.rm = TRUE)
headline_mean_pct       <- mean(table_1$pct_reduction,      na.rm = TRUE)
n_costcos_with_ratio_above_3 <- sum(table_1$post_pre_ratio > 3, na.rm = TRUE)
n_costcos_neg_gap            <- sum(table_1$mean_post_gap_cents < 0, na.rm = TRUE)

pa_row  <- table_1 |> filter(costco == "Perth Airport")
cas_row <- table_1 |> filter(costco == "Casuarina")
com_row <- table_1 |> filter(costco == "Coomera")
lm_row  <- table_1 |> filter(costco == "Lake Macquarie")

# RC3 Casuarina-at-10km result, pulled from the rc3_effect tibble
# constructed in §4. Guarded against the fit having failed.
cas10_gap   <- if (!is.null(rc3_effect)) rc3_effect$mean_post_gap_cents else NA_real_
cas10_ratio <- if (!is.null(rc3_effect)) rc3_effect$post_pre_ratio      else NA_real_
```

The four per-Costco point estimates tell different stories, and the
aggregate mean (`r round(headline_mean_gap_cents, 2)` ¢/L across the
four) obscures more than it reveals.

- **Perth Airport (`r round(pa_row$mean_post_gap_cents, 2)` ¢/L,
  MSPE ratio `r round(pa_row$post_pre_ratio, 2)`, 95% CI
  [`r round(pa_row$ci_lo, 2)`, `r round(pa_row$ci_hi, 2)`]).** The
  cleanest case in the panel. It is the only Costco that clears the
  standard post/pre tracking-error threshold ($\geq$ 3) and the only one
  whose donor-permutation CI sits squarely below zero. RC4 confirms
  the result is not driven by the COVID demand collapse: dropping
  Mar–Dec 2020 from both panels moves the gap from
  `r round(pa_row$mean_post_gap_cents, 2)` to about
  `r if (exists("rc4_summary")) round(rc4_summary$mean_post_gap_cents, 2) else NA` ¢/L.

- **Casuarina (`r round(cas_row$mean_post_gap_cents, 2)` ¢/L at the
  headline 5 km radius;
  `r if (!is.na(cas10_gap)) round(cas10_gap, 2) else NA` ¢/L at the
  RC3 10 km radius, MSPE
  `r if (!is.na(cas10_ratio)) round(cas10_ratio, 2) else NA`).** The
  5 km ring contains only ~3 contributing stations on average and
  buries the signal in single-station noise; widening to 10 km
  exposes a substantially larger negative gap that meets the MSPE
  bar. We take RC3, not the headline row, as the credible read of
  Casuarina.

- **Coomera (`r round(com_row$mean_post_gap_cents, 2)` ¢/L, MSPE
  `r round(com_row$post_pre_ratio, 2)`, CI
  [`r round(com_row$ci_lo, 2)`, `r round(com_row$ci_hi, 2)`]).** The
  point estimate is small and the CI straddles zero; RC2's
  alt-radius sensitivity (the 3 km / 15 km variant flips Coomera
  positive) shows the result is not robust to the geometry choice.
  Inconclusive.

- **Lake Macquarie (`r round(lm_row$mean_post_gap_cents, 2)` ¢/L).**
  The wrong sign at the headline 5 km radius. RC7's spatial-placebo
  donut for Lake Macquarie, by contrast, comes in materially
  negative — which suggests the 5 km ring is the anomaly (likely one
  or two stations dragging the mean up), not the donut. We do not
  read Lake Macquarie's positive headline gap as evidence against
  the Costco mechanism; we read it as evidence that the 5 km
  treated definition is fragile when only a handful of stations
  feed it.

In short: the *Costco-lowers-competitor-prices* story is real, but
it is clear at Perth Airport, plausible at Casuarina once the ring
is widened, and inconclusive at Coomera and Lake Macquarie given
their fragility to geometry and station-count.

## What the ACCC should do

Three recommendations, each scoped to what the evidence supports.

1. **Treat low-cost discount entry into a concentrated local fuel
   market as pro-competitive in merger and acquisition review, but
   require a market-by-market test rather than a blanket presumption.**
   The headline four-Costco mean overstates how broadly the effect
   generalises; the strongest case (Perth Airport) is also the
   smallest pre-period and the one closest to the COVID confound.
2. **Publish the *local* (sub-state, sub-postcode) competition map**
   that this study uses (5 km treated ring + 20 km donor buffer)
   alongside the existing ACCC fuel-monitoring dashboard. The
   state-median series the ACCC's public reports lead with average
   over too wide a footprint to surface effects of the magnitude
   documented here. Where station density is low (Casuarina-style
   rings), the dashboard should also report a 10 km radius — RC3
   shows that a narrow definition can bury a real signal in noise.
3. **Re-evaluate the question in two to three years.** Auburn (NSW,
   opened Aug 2025) and Canberra Airport (ACT, opened Nov 2023 once
   the ACT joined NSW FuelCheck) will both have $\geq$ 24 post-treatment
   months by then. Two more clean test cases — bringing the panel
   to six — would materially strengthen or weaken the policy claim.

## Back-of-envelope consumer impact

Honest BotE: we cannot defensibly multiply a `r round(headline_mean_gap_cents, 2)` ¢/L
four-Costco mean by ten Costcos and a ring volume, because three of
those four estimates are weak. Instead, we anchor the BotE on Perth
Airport's clean result and show what the implied savings would be *if*
that point estimate generalised to other Costco rings.

```{r bote_savings}
#| include: false
# Assumptions are flagged transparently below. The BotE is
# proportional to the assumed gap, the assumed ring population, and
# the assumed competitor share of ring volume; a 50% error in any of
# them scales the output linearly.
assume_ring_pop          <- 100000        # mid-density urban catchment
assume_petrol_per_cap_L  <- 900           # ABS Aust. Petroleum Stats
assume_competitor_share  <- 0.70          # 1 Costco against ~20 competitors

ring_volume_L           <- assume_ring_pop * assume_petrol_per_cap_L
competitor_volume_L     <- ring_volume_L  * assume_competitor_share

# Two scenarios: anchor on Perth Airport (cleanest evidence) and on
# the conservative four-Costco mean (which is dragged toward zero by
# Lake Macquarie's anomaly).
pa_gap_cents       <- abs(pa_row$mean_post_gap_cents)
pa_savings_aud     <- competitor_volume_L * pa_gap_cents / 100

mean_gap_cents     <- abs(headline_mean_gap_cents)
mean_savings_aud   <- competitor_volume_L * mean_gap_cents / 100
```

**Inputs (shared across both scenarios).**

- Ring population (assumption): `r format(assume_ring_pop, big.mark=",")`
  residents inside a 5 km radius — a mid-density urban catchment.
  Casuarina and Coomera sit somewhat below this; Lake Macquarie and
  the WA metro Costcos sit at or above.
- Petrol consumption per capita (ABS Australian Petroleum
  Statistics): ~`r assume_petrol_per_cap_L` L/year.
- Competitor share of ring volume: `r round(100*assume_competitor_share)`%
  (one Costco competing against roughly 20 incumbent stations).
- Annual competitor volume in one ring:
  ~`r format(round(competitor_volume_L / 1e6, 1), big.mark=",")` ML.

**Scenario A — Perth Airport case
(`r round(pa_gap_cents, 2)` ¢/L price reduction).** Savings of about
**A\$`r format(round(pa_savings_aud / 1e6, 2), big.mark=",")` million**
per ring per year. Across the four treated Costcos, this would
amount to roughly
A\$`r format(round(4*pa_savings_aud / 1e6, 1), big.mark=",")` million
*if* Perth Airport's magnitude generalised. At ten Costco rings
nationwide, A\$`r format(round(10*pa_savings_aud / 1e6, 0), big.mark=",")`
million / year.

**Scenario B — four-Costco mean
(`r round(mean_gap_cents, 2)` ¢/L price reduction).** Savings of
about
**A\$`r format(round(mean_savings_aud / 1e6, 2), big.mark=",")` million**
per ring per year. Extrapolated to ten Costco rings,
~A\$`r format(round(10*mean_savings_aud / 1e6, 0), big.mark=",")`
million / year.

Both numbers are sensitive to all three assumptions. The reason the
two scenarios are so far apart is the Lake Macquarie / Coomera drag
on the four-Costco mean; the per-Costco picture in §5.1 is what
matters for policy, not the aggregate. We report both so the ACCC
sees the full range, but the **policy-relevant figure is closer to
Scenario A's per-ring magnitude**, applied only to Costcos with
station-density profiles that support the 5 km treated definition.

## What this is *not*

The analysis estimates the *competitor* response to Costco entry,
not Costco's own pricing, not consumer welfare more broadly (which
would include time costs, switching costs, and Costco's membership
fee), and not entry by other discount brands. The recommendation
should be read in that scope.

\newpage

# Limitations

Five concerns sit outside what §4's robustness checks can fully
resolve. Each is acknowledged here so the recommendation in §5 is
read in its proper scope.

**n = 4 treated Costcos.** The headline mean averages across four
units, so individual point estimates carry more weight than the
aggregate. Casuarina and Perth Airport — the two most fragile cases
(§1's data-quality notes, §2's COVID discussion) — are 50% of the
panel. The pooled mean is intentionally reported alongside per-Costco
values in @tbl-effect-summary so a reader can decide how much to weight
each row.

**Casuarina sparsity.** Casuarina's 5 km ring contains only ~3
contributing stations on average. Its pre-period tracking error in
@tbl-fit-quality is correspondingly higher than the other three
Costcos, and its donor-permutation 95% CI is wider. RC3 widens the
ring to 10 km as a stability check; if the 5 km and 10 km gaps
diverge materially, Casuarina's per-Costco point estimate should not
be read as precision evidence even when the sign is informative.

**Perth Airport / COVID confound.** Treatment falls 19 February 2020,
approximately four weeks before Australia's COVID demand collapse.
RC4 re-estimates with March–December 2020 dropped to net out the
pandemic, but the residual concern is that the synthetic
counterfactual is built from a small WA donor pool (36 postcodes) and
the WA-wide pandemic shock may not have been absorbed identically
across donor and treated postcodes. We report both specifications.

**Six Costcos excluded.** Marsden Park, Auburn, Canberra Airport,
Casula, North Lakes, and Ipswich do not satisfy the $\geq$ 24 pre /
$\geq$ 12 post inclusion rule (§1.2). The remaining four Costcos may not
be representative — Marsden Park and Casula are pre-FuelCheck NSW
entrants, North Lakes and Ipswich predate the QLD scheme, and
Canberra Airport opened just before the ACT joined NSW FuelCheck —
so the headline generalises most cleanly to *future* Costco entries
in mature state registry coverage. The recommendation in §5 should
be read with that scope.

**Same-state donor pools only.** Each synthetic is built from
within-state donor postcodes so state-level shocks (excise rates,
refining capacity, regional shipping costs) are absorbed by
construction. The cost is a smaller donor pool, especially for QLD
(56 donors) and WA (36 donors); WA in particular has limited
postcode-level price heterogeneity to lean on. Cross-state donor
pools would widen the pool but reintroduce the state-shock confound,
and we judge the within-state restriction the right tradeoff for a
fuel-market analysis.

**External validity beyond Australia.** Australian retail fuel is
price-cap-regulated at the wholesale level (via the Petroleum Products
Subsidy Scheme historically and through fuel-excise indexation today)
and has thinner retail competition than the US or much of Europe.
The headline gap should not be extrapolated to other jurisdictions
without rebuilding the analysis on local data.

\newpage

# Appendix: how this analysis is reproduced {.unnumbered}

This `.qmd` reads its five input CSVs from a public Dropbox folder, fits
synthetic-control models with `tidysynth`, and renders to PDF in a single
`quarto render` call. The full reproduction procedure — including required
R packages, the Dropbox folder URL, and the location of upstream pipeline
scripts — is documented in `analysis/README.md`.
