---
title: "Does Costco Entry Lower Competitor Fuel Prices in Australia?"
subtitle: "A Synthetic-Control Study for the ACCC"
author: "MGT159 Group Project — Costco Australia"
date: today
format:
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    toc: true
    toc-depth: 2
    number-sections: true
    fig-pos: "H"
    fig-align: center
    fig-width: 7
    fig-height: 4.5
    colorlinks: true
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    dev: "pdf"
    out.width: "100%"
    cache: false
---

```{r setup}
#| cache: false
#| include: false
source("_sc_helpers.R")

library(tidyverse)
library(tidysynth)
library(lubridate)
library(kableExtra)
library(patchwork)
library(scales)
library(glue)

# Reproducibility — the five analysis-ready CSVs live in one public Dropbox
# folder. Appending dl=1 returns the folder contents as a ZIP that we
# download once per session, unzip into tempdir, and read from there.
DROPBOX_FOLDER_URL <- "https://www.dropbox.com/scl/fo/g7lhsenilo4lza6dwnukq/AMIiPcOy1uysFMJtSC3eqrg?rlkey=ozb016wrtnrg9qj9yw789xkyg&st=73xiohe7&dl=1"

cache_dir <- file.path(tempdir(), "costco_sc_inputs")
dir.create(cache_dir, showWarnings = FALSE, recursive = TRUE)

required <- c("treated_units.csv", "donor_pool.csv", "treated_metadata.csv",
              "donor_metadata.csv", "state_median_monthly.csv")

if (!all(file.exists(file.path(cache_dir, required)))) {
  zip_path <- file.path(cache_dir, "inputs.zip")
  download.file(DROPBOX_FOLDER_URL, zip_path, mode = "wb", quiet = TRUE)
  unzip(zip_path, exdir = cache_dir, junkpaths = TRUE)
  missing <- required[!file.exists(file.path(cache_dir, required))]
  if (length(missing) > 0) {
    stop("After unzip, missing files: ", paste(missing, collapse = ", "),
         ". Check the Dropbox folder contents.")
  }
}

treated_units    <- read_csv(file.path(cache_dir, "treated_units.csv"),
                             show_col_types = FALSE)
donor_pool       <- read_csv(file.path(cache_dir, "donor_pool.csv"),
                             show_col_types = FALSE)
treated_metadata <- read_csv(file.path(cache_dir, "treated_metadata.csv"),
                             show_col_types = FALSE)
donor_metadata   <- read_csv(file.path(cache_dir, "donor_metadata.csv"),
                             show_col_types = FALSE)
state_median     <- read_csv(file.path(cache_dir, "state_median_monthly.csv"),
                             show_col_types = FALSE)

# State display order — alphabetical for figures, fixed for tables.
state_levels <- c("NSW", "QLD", "WA")

# Treated-Costco order (south-to-north-ish, matches plan-of-attack figures).
costco_order <- c("Perth Airport", "Casuarina", "Coomera", "Lake Macquarie")
```

\newpage

# Question and motivation

We advise the **Australian Competition and Consumer Commission (ACCC)**,
Australia's national competition regulator. The ACCC publishes regular
fuel-price monitoring reports and has flagged ongoing concerns about market
concentration in Australian fuel retailing. One of the ACCC's main tools for
improving competition in concentrated retail markets is reducing entry barriers
for discount entrants — but whether such entrants actually discipline
incumbent pricing is an empirical question, not a foregone conclusion.

We ask:

> **Does the entry of a Costco gas station into an Australian local market
> cause nearby competitor stations to lower their retail fuel prices?**

This is a causal question. We answer it using **synthetic control**, fit
separately for each of the four Australian Costco fuel stations that satisfy
our pre- and post-treatment data-availability rules: Perth Airport (WA, Feb 2020),
Casuarina (WA, Nov 2022), Lake Macquarie (NSW, May 2022), and Coomera (QLD,
May 2023). The complete plan, including identifying assumptions, sample
construction, and the seven pre-committed robustness checks, is in
`deliverables/plan_of_attack_section_{1,2,3}.pdf`. This document is the
*execution* of that plan.

\newpage

# Section 1: Raw Data Description

## Data sources

We combine three publicly accessible Australian state fuel-price registries —
**NSW FuelCheck** (data.nsw.gov.au), the **Queensland Fuel Price Reporting
Scheme** (data.qld.gov.au), and **WA FuelWatch** (data.wa.gov.au) — with a
hand-collected list of Costco fuel-station opening dates. The three registries
together yield approximately **17.4 million station-day price records across
6,084 unique stations between December 2016 and April 2026**. Each registry
mandates public price reporting by every retail fuel station in the state.

```{r tbl-registries}
tibble::tribble(
  ~Registry,        ~State, ~Coverage,                 ~`Raw records`, ~`Unique stations`,
  "NSW FuelCheck",  "NSW",  "Dec 2016 – Jan 2026",     "4,401,971",    "4,813",
  "QLD FPRS",       "QLD",  "Dec 2018 – Dec 2025",     "4,970,650",    "1,883",
  "WA FuelWatch",   "WA",   "Jan 2018 – Apr 2026",     "8,047,153",    "1,423"
) |>
  kbl(booktabs = TRUE, align = "lllrr",
      caption = "State fuel-price registries used in this analysis.") |>
  kable_styling(latex_options = c("striped", "hold_position"))
```

NSW FuelCheck and QLD FPRS record price-change events (a row per
station × fuel type × price-update timestamp); WA FuelWatch publishes daily
snapshots (a row per station × fuel type × calendar day). All three are
harmonised in `scripts/synthetic_control_input/build_sc_inputs.py` to a
common station × month panel of mean unleaded price in cents per litre.
Detailed field-level schemas appear in
`deliverables/plan_of_attack_section_1.pdf` §1.1–1.3.

## Treated Costco fuel stations

Ten Costco fuel stations operate (or recently opened) in Australia. Four
satisfy the synthetic-control inclusion rule (≥24 pre-treatment months and
≥12 post-treatment months in the relevant state's registry) and form the
treated panel; the remaining six are excluded from formal estimation but
appear here as descriptive context.

```{r tbl-costco-status}
treated_rows <- treated_metadata |>
  transmute(
    Costco = costco_key,
    State  = state,
    `Treatment date` = format(as.Date(treatment_date), "%b %Y"),
    `Pre months`  = n_pre_months,
    `Post months` = n_post_months,
    Status = "Treated"
  )

# Six excluded Costcos — from plan-of-attack §1.4, hand-collected metadata.
excluded_rows <- tibble::tribble(
  ~Costco,            ~State, ~`Treatment date`, ~`Pre months`, ~`Post months`, ~Status,
  "Marsden Park",     "NSW",  "Jul 2017",        11L,           105L,           "Excluded (pre too short)",
  "Auburn",           "NSW",  "Aug 2025",        107L,            8L,           "Excluded (post too short)",
  "Canberra Airport", "ACT",  "Nov 2023",         0L,            29L,           "Excluded (no pre)",
  "Casula",           "NSW",  "Dec 2016",         4L,           113L,           "Excluded (pre-registry)",
  "North Lakes",      "QLD",  "Jan 2019",         1L,            89L,           "Excluded (pre-registry)",
  "Ipswich",          "QLD",  "Mar 2019",         3L,            87L,           "Excluded (pre-registry)"
)

bind_rows(treated_rows, excluded_rows) |>
  kbl(booktabs = TRUE, align = "llrrrl",
      caption = "Australian Costco fuel stations by treatment status.") |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9)
```

The four treated Costcos give us **staggered treatment timing across three
states** (Feb 2020, May 2022, Nov 2022, May 2023), which is the variation
the synthetic-control design exploits.

## The analysis panel

After ingest, geo-coding (Costco distance), the regular-unleaded filter,
sentinel-price drops, and two-stage station × month → ring/postcode × month
aggregation, our analysis panel contains two pieces:

```{r tbl-panel-summary}
panel_summary <- tibble(
  Component = c(
    "Treated 5 km rings × month",
    "Donor postcodes × month",
    "Combined analysis panel"
  ),
  Rows = c(nrow(treated_units),
           nrow(donor_pool),
           nrow(treated_units) + nrow(donor_pool)),
  Units = c(
    glue("{n_distinct(treated_units$costco_key)} treated Costcos"),
    glue("{n_distinct(donor_pool$postcode)} donor postcodes",
         " (NSW = {sum(donor_metadata$state == 'NSW')},",
         " QLD = {sum(donor_metadata$state == 'QLD')},",
         " WA = {sum(donor_metadata$state == 'WA')})"),
    "treated + donor"
  )
)

panel_summary |>
  kbl(booktabs = TRUE,
      format.args = list(big.mark = ","),
      caption = "Analysis-ready panel sizes after sample-construction filters.") |>
  kable_styling(latex_options = c("striped", "hold_position"))
```

The sample-construction funnel (raw rows → aggregated cells → final panel)
is reported in §2.5 as documentary text, since reconstructing it inside this
.qmd would require pulling the full 1.86 GB of raw API archives.

## Distribution of monthly mean unleaded price

```{r fig-price-histogram}
#| fig-cap: "Distribution of monthly mean unleaded price across the combined SC analysis panel (treated 5 km rings + donor postcodes). The distribution is approximately bimodal: the lower mode reflects the 2018–2020 price regime; the higher mode reflects the post-2022 elevated price era following the Russia/Ukraine energy shock."
#| fig-height: 4

panel_combined <- bind_rows(
  treated_units |> transmute(date = as.Date(date), mean_price_cents,
                             segment = "Treated 5 km ring"),
  donor_pool    |> transmute(date = as.Date(date), mean_price_cents,
                             segment = "Donor postcode")
) |>
  filter(!is.na(mean_price_cents))

panel_mean   <- mean(panel_combined$mean_price_cents, na.rm = TRUE)
panel_median <- median(panel_combined$mean_price_cents, na.rm = TRUE)

ggplot(panel_combined, aes(x = mean_price_cents)) +
  geom_histogram(binwidth = 3, fill = "#4C72B0", color = "white", alpha = 0.9) +
  geom_vline(xintercept = panel_mean,   linetype = "dashed",
             color = "#C44E52", linewidth = 0.6) +
  geom_vline(xintercept = panel_median, linetype = "dashed",
             color = "#DD8452", linewidth = 0.6) +
  annotate("label", x = panel_mean,   y = Inf, vjust = 1.5,
           label = glue("mean = {round(panel_mean,   1)} ¢/L"),
           color = "#C44E52", size = 3, label.size = 0) +
  annotate("label", x = panel_median, y = Inf, vjust = 3.5,
           label = glue("median = {round(panel_median, 1)} ¢/L"),
           color = "#DD8452", size = 3, label.size = 0) +
  labs(x = "Postcode-month mean unleaded price (¢/L)",
       y = "Frequency",
       title = glue("Distribution of monthly mean unleaded price (n = {comma(nrow(panel_combined))})")) +
  theme_minimal(base_size = 11)
```

## Monthly median unleaded price by state

```{r fig-state-median-time}
#| fig-cap: "Median of postcode-monthly mean unleaded prices over time, by state. All three states track each other closely: prices fall in 2018–2020 (peaking with the COVID demand collapse in early 2020), rise sharply through 2022 (Russia/Ukraine energy shock), dip during the Mar–Sep 2022 federal fuel-excise cut, and remain elevated through 2026. State-level co-movement supports our use of within-state donor pools to absorb common national shocks."
#| fig-height: 4

state_median |>
  mutate(date = as.Date(date),
         state = factor(state, levels = state_levels)) |>
  ggplot(aes(x = date, y = median_price_cents, color = state)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 0.8, alpha = 0.7) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_color_manual(values = c(NSW = "#4C72B0", QLD = "#DD8452", WA = "#55A868"),
                      name = "State") +
  labs(x = NULL, y = "Median monthly unleaded price (¢/L)",
       title = "Monthly median unleaded price by state, 2017-2026") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "top")
```

## Treated 5 km rings versus state medians

```{r fig-treated-vs-state}
#| fig-cap: "Treated 5 km ring mean unleaded price (colored solid) versus the state median across all reporting postcodes (gray dashed) for each of the four treated Costcos. The vertical red dashed line marks the validated treatment date. Shaded gray: the COVID demand collapse (Mar–Sep 2020); dotted vertical: Russia's invasion of Ukraine (24 Feb 2022); pale-blue shading: the AU federal fuel-excise cut (30 Mar – 28 Sep 2022). The state-median overlay is a descriptive level reference, not the synthetic-control counterfactual."
#| fig-height: 7
#| fig-width: 8

# Shock annotations (Plan of Attack Section 1)
covid_start  <- as.Date("2020-03-01")
covid_end    <- as.Date("2020-09-30")
ukraine      <- as.Date("2022-02-24")
excise_start <- as.Date("2022-03-30")
excise_end   <- as.Date("2022-09-28")

state_colors <- c(NSW = "#C44E52", QLD = "#4C72B0",
                   WA  = "#55A868")

make_panel <- function(costco_name) {
  meta <- treated_metadata |> filter(costco_key == costco_name)
  tr   <- treated_units    |> filter(costco_key == costco_name) |>
    mutate(date = as.Date(date))
  sm   <- state_median     |> filter(state == meta$state) |>
    mutate(date = as.Date(date))

  ggplot() +
    annotate("rect", xmin = covid_start, xmax = covid_end,
             ymin = -Inf, ymax = Inf, fill = "gray80", alpha = 0.5) +
    annotate("rect", xmin = excise_start, xmax = excise_end,
             ymin = -Inf, ymax = Inf, fill = "#B0CFE0", alpha = 0.5) +
    geom_vline(xintercept = ukraine, linetype = "dotted",
               color = "gray30", linewidth = 0.4) +
    geom_line(data = sm, aes(x = date, y = median_price_cents),
              color = "gray50", linetype = "dashed", linewidth = 0.5) +
    geom_line(data = tr, aes(x = date, y = mean_price_cents),
              color = state_colors[[meta$state]], linewidth = 0.7) +
    geom_vline(xintercept = as.Date(meta$treatment_date),
               linetype = "dashed", color = "#C44E52", linewidth = 0.6) +
    scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
    labs(x = NULL, y = "¢/L",
         title = glue("{costco_name} ({meta$state}): 5 km ring vs {meta$state} state median"),
         subtitle = glue("opens {format(as.Date(meta$treatment_date), '%d %b %Y')}")) +
    theme_minimal(base_size = 9) +
    theme(plot.title = element_text(size = 10, face = "bold"),
          plot.subtitle = element_text(size = 8, color = "gray40"))
}

panels <- map(costco_order, make_panel)
wrap_plots(panels, ncol = 1)
```

## Summary statistics

```{r tbl-summary-stats}
price_stats <- panel_combined |>
  summarize(
    n      = n(),
    Mean   = mean(mean_price_cents, na.rm = TRUE),
    SD     = sd(mean_price_cents, na.rm = TRUE),
    Min    = min(mean_price_cents, na.rm = TRUE),
    p25    = quantile(mean_price_cents, 0.25, na.rm = TRUE),
    Median = median(mean_price_cents, na.rm = TRUE),
    p75    = quantile(mean_price_cents, 0.75, na.rm = TRUE),
    Max    = max(mean_price_cents, na.rm = TRUE)
  ) |>
  mutate(Variable = "Monthly mean unleaded price (¢/L)", .before = 1)

dist_stats <- donor_metadata |>
  summarize(
    n      = n(),
    Mean   = mean(min_dist_any_costco_km, na.rm = TRUE),
    SD     = sd(min_dist_any_costco_km, na.rm = TRUE),
    Min    = min(min_dist_any_costco_km, na.rm = TRUE),
    p25    = quantile(min_dist_any_costco_km, 0.25, na.rm = TRUE),
    Median = median(min_dist_any_costco_km, na.rm = TRUE),
    p75    = quantile(min_dist_any_costco_km, 0.75, na.rm = TRUE),
    Max    = max(min_dist_any_costco_km, na.rm = TRUE)
  ) |>
  mutate(Variable = "Donor-postcode distance to nearest Costco (km)", .before = 1)

bind_rows(price_stats, dist_stats) |>
  mutate(across(c(Mean, SD, Min, p25, Median, p75, Max), \(x) round(x, 1))) |>
  kbl(booktabs = TRUE, align = "lrrrrrrrr",
      format.args = list(big.mark = ","),
      caption = "Summary statistics on the SC analysis panel.") |>
  kable_styling(latex_options = c("striped", "hold_position", "scale_down"))
```

The mean unleaded price row matches the combined treated + donor sample. The
distance row is computed at the donor-postcode level (one row per postcode);
the plan-of-attack Section 1 reports the analogous statistic at the station
level (n = 5,980, mean ≈ 205 km), so the numbers differ in denominator but
agree in shape: most non-Costco stations sit very far from any Costco.

## Data quality and known limitations

- **Price-unit harmonisation.** NSW and WA prices are recorded in cents/L;
  QLD records prices in tenths of cents (e.g. `1899` = 189.9 ¢/L). The build
  pipeline applies the appropriate scaling per registry so the analysis
  panel is uniform in cents/L.
- **Missing coordinates in historical NSW and WA files.** Both registries
  publish historical files without lat/lng; we resolve this with a one-time
  pull from each state's live API, joined to historical records by
  (normalized name, postcode). Stations that closed or rebranded before the
  live-API snapshot fail the join and drop out (52.3% of NSW historical
  stations, 47.3% of WA). The loss falls almost entirely on the donor pool;
  all four treated Costcos join cleanly.
- **NSW Costco rename in May 2022.** Every NSW Costco station was renamed
  in the registry from e.g. `Costco Marsden Park` to
  `Costco Marsden Park (Members only)` on 2 May 2022. The build pipeline
  strips the suffix before matching.
- **Sentinel-price filter.** Quotes outside 50–500 ¢/L are dropped before
  aggregation; this removes obvious out-of-stock placeholders like `9999`
  or `0.01`. After the filter, no postcode-month in the analysis panel
  falls outside 92–259 ¢/L (well inside the plausible Australian range).
- **Casuarina sparse ring.** Casuarina's 5 km ring contains only about
  three contributing stations on average — a documented limitation flagged
  in advance. Robustness Check 3 (Casuarina at 10 km) probes whether this
  noise drives the headline.
- **Perth Airport opens four weeks before COVID.** Its early post-treatment
  months are dominated by pandemic-driven price moves; Robustness Check 4
  (drop Mar–Dec 2020) probes whether this confounds the headline.

\newpage

# Section 2: Sample Construction

Every choice between raw registry rows and the analysis panel deserves
justification — the headline result is a causal claim about prices, so
opacity at the construction stage would undercut everything downstream.
The subsections below mirror Plan-of-Attack §2(a)–(g) in order. This
section documents the 5 km treated ring / 20 km donor buffer geometry;
alternate geometries appear in §4 Robustness Checks.

## Sample window

The pre-period for each treated Costco runs from the start of its
state's price registry — December 2016 for NSW, January 2018 for WA,
December 2018 for QLD — through the day before that Costco's validated
treatment date (§2.4). The post-period runs from the treatment date
through April 2026, the most recent registry pull. Per-Costco coverage:

```{r tbl-coverage}
treated_metadata |>
  transmute(
    Costco           = costco_key,
    State            = state,
    `Treatment date` = format(as.Date(treatment_date), "%b %Y"),
    `Pre months`     = n_pre_months,
    `Post months`    = n_post_months
  ) |>
  kbl(booktabs = TRUE, align = "llrrr",
      caption = "Pre- and post-treatment coverage per treated Costco.") |>
  kable_styling(latex_options = c("striped", "hold_position"))
```

Synthetic control fits its weights on the entire pre-period *trajectory*
of the dependent variable, not just on a level. More pre-period
observations tighten identification rather than dilute it — the opposite
of the DID logic in which long pre-periods can introduce structural
breaks into the parallel-trends assumption. Same-state donors experience
the same world-shifts (oil-price regime changes, vehicle fleet turnover,
federal excise adjustments) over the same window, so common shocks are
absorbed into the synthetic by construction. We use as much pre-period
as the data allow.

**COVID.** The 2020 demand collapse and recovery is the only structural
break large enough to warrant explicit handling. Three of the four
treated Costcos opened well after the COVID disruption (Lake Macquarie
May 2022, Casuarina November 2022, Coomera May 2023); for those three
the COVID period sits entirely inside the pre-period and is matched by
the synthetic by construction. **Perth Airport (treated 19 February
2020) opens approximately four weeks before Australia's COVID demand
collapse begins; this is the most fragile case in the analysis.** Its
pre-period is entirely pre-COVID, but its early post-treatment months
are dominated by pandemic-driven price movements rather than competitive
response to Costco's entry. The synthetic counterfactual is built from
WA donor postcodes that experienced the same pandemic shock, so the
post-treatment gap nets out COVID provided the pre-period weights track
the treated trajectory through January 2020. §4 Robustness Check 4
re-estimates Perth Airport with March–December 2020 dropped from both
panels.

## Other sample restrictions

**Treated unit.** For each treated Costco, the treated unit is the
*local market* defined as all non-Costco retail stations within 5 km of
the Costco's coordinates, summarized each month as the simple mean of
those stations' monthly mean unleaded prices. Costco's own price quotes
are excluded from the treated mean — the research question concerns
competitor response, not Costco's own pricing.

**Donor pool.** Each donor unit is a 4-digit Australian postcode,
summarized the same way (monthly mean of non-Costco unleaded price
across all stations within the postcode). Postcodes are the natural
geographic unit of all three state registries and are consistent across
NSW, QLD, and WA. A postcode is excluded from the donor pool if any of
its stations lie within 20 km of any Costco fuel station (treated,
excluded, or otherwise), so donor postcodes are unambiguously outside
any plausible Costco competitive footprint.

**Asymmetric radii (5 km treated, 20 km donor buffer).** The economics
literature on retail fuel-price competition (Houde 2012; Eckert 2013)
finds price effects concentrated within roughly 1–5 miles. 5 km
(~3.1 mi) sits at the conservative end of that range — small enough to
capture genuinely competing stations but not so small as to drop most
of the local market. The 20 km donor buffer is intentionally a 4×
multiple, providing a wide exclusion donut so that no donor postcode is
partly affected by a Costco that isn't the one we're studying. The
asymmetry biases the design *against* finding an effect: a narrow
treated definition shrinks the treated population to the most exposed
stations, and the wide buffer shrinks the donor pool to truly untreated
postcodes.

**Outcome variable.** Regular unleaded — `FuelCode` in
`{U91, E10, P91}` for NSW; `Fuel_Type` in `{Unleaded 91, e10}` for QLD;
`PRODUCT_DESCRIPTION = ULP` for WA. Unleaded is the highest-volume
retail product in Australia, comparable across the three state schemes,
and Costco's headline competitive product.

**Donor-pool measurement-quality filters.** A postcode is additionally
excluded if it averages fewer than 3 unique stations per month (monthly
means based on 1–2 stations are too noisy to serve as credible
counterfactual building blocks) or has fewer than 24 months of any
observed coverage (weights cannot be reliably fit on a short
pre-period).

**Treated-Costco inclusion rule.** A Costco fuel station is treated as
a synthetic-control unit if and only if its registry history contains
both **at least 24 pre-treatment months** *and* **at least 12
post-treatment months** in the relevant state's price registry. Four
of Australia's ten Costco fuel stations satisfy both conditions; six
do not and appear as descriptive context only (§1.2).

```{r tbl-thresholds}
tibble::tribble(
  ~Threshold,               ~Value,         ~Justification,                                                                  ~`Robustness check (§4)`,
  "Treated radius",         "≤ 5 km",       "Conservative end of IO-literature retail fuel-price spillover range (1–5 mi)",   "Re-run at 3 km and 8 km",
  "Donor buffer",           "> 20 km",      "4× the treated radius; keeps donors fully outside any Costco's competitive zone","Re-run at 15 km and 30 km",
  "Donor stations / month", "≥ 3",          "Below 3, monthly mean has high sampling variance",                               "Re-run at ≥ 2 and ≥ 5",
  "Donor coverage",         "≥ 24 months",  "Minimum pre-period length for stable SC weight fitting",                         "Re-run at ≥ 36 months",
  "Costco pre-period",      "≥ 24 months",  "Same logic, applied to treated",                                                 "Not relaxed (would re-introduce excluded Costcos)",
  "Costco post-period",     "≥ 12 months",  "Minimum to estimate a sustained-effect gap",                                     "Not relaxed",
  "Plausible price range",  "50–500 ¢/L",   "Sentinel-value floor and ceiling; AU retail unleaded lives in 80–280 ¢/L",       "None; not an analytic threshold"
) |>
  kbl(booktabs = TRUE, align = "llll",
      caption = "Threshold-and-defense summary for the SC sample.") |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9) |>
  column_spec(1, width = "2.8cm") |>
  column_spec(2, width = "2cm") |>
  column_spec(3, width = "5.7cm") |>
  column_spec(4, width = "4.5cm")
```

## Dependent variable integrity

**Construction.** The dependent variable is the monthly mean unleaded
retail price (cents per litre) at the treated 5 km ring or the donor
postcode; see §2.5 for the two-stage construction. @fig-price-histogram shows
the post-aggregation distribution (n = `r comma(nrow(panel_combined))`
postcode-months, range 92–259 ¢/L, approximately bimodal across the
pre-2022 / post-2022 price-regime divide).

**Sentinel-value filter.** Individual price quotes outside 50–500 ¢/L
are dropped before aggregation. This is a data-integrity filter, not
winsorisation: it removes obvious registry artifacts (out-of-stock
placeholder values like `9999` or `0.01`) rather than legitimate tail
observations.

**Missing months.** A postcode-month or treated ring-month with zero
contributing station observations is left as missing (NA), never imputed
as zero, never carried forward. Synthetic-control fitting handles
missing months by ignoring them in the pre-period objective. §1's
data-quality notes report that the largest run of missing months in any
treated panel is three (Lake Macquarie), not material.

**No winsorisation in the main specification.** Two-stage monthly
aggregation smooths individual-quote outliers: an extreme single quote
at one station gets averaged with the rest of that station's month, and
then with the rest of the ring or postcode. If a §4 robustness check
exposes a donor-postcode month driving an estimate, we winsorise the
analysis panel at the 99th percentile and report both specifications
side by side.

## Independent variable integrity

**Treatment indicator.** A binary post-opening flag is constructed per
treated Costco by merging the hand-collected treatment-date table onto
the (Costco × month) panel: 1 from the treatment month forward, 0
before.

**Treatment-date validation.** Registry first-appearance can lag the
true fuel-station opening if a station begins selling fuel before
appearing in the state's reporting feed. Each of the four treated
Costcos' registry-first-appearance dates was cross-checked against
Costco Australia corporate announcements, trade press, and news
coverage. The corroborating source must specifically reference the
*fuel station* rather than the warehouse; Costco Australia warehouses
and fuel stations regularly open weeks apart. If a fuel-station-specific
announced opening date differs from registry first-appearance by more
than 30 days, the announced date supersedes.

```{r tbl-treatment-dates}
tibble::tribble(
  ~Costco,           ~`Registry first-appearance`, ~`Fuel-station-specific corroboration`,                                                  ~`Delta`,    ~`Treatment date used`,
  "Coomera",         "23 May 2023",                "23 May 2023 (Retail World; explicit fuel-station opening)",                              "0 d",       "23 May 2023 (registry)",
  "Casuarina",       "11 Nov 2022",                "None; trade press only reports the warehouse opening (7 Dec 2022)",                      "n/a",       "11 Nov 2022 (registry)",
  "Perth Airport",   "11 Apr 2020",                "19 Feb 2020 (ACAPMAg: petrol station selling at \\$1.18/L; warehouse opens 19 Mar 2020)", "-52 d",     "\\textbf{19 Feb 2020 (announced)}",
  "Lake Macquarie",  "2 May 2022",                 "None; no fuel-station-specific opening date in public sources",                          "n/a",       "2 May 2022 (registry)"
) |>
  kbl(booktabs = TRUE, align = "lllll", escape = FALSE,
      caption = "Validated treatment dates: registry first-appearance cross-checked against fuel-station-specific corroboration. Perth Airport is the only revision.") |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9) |>
  column_spec(1, width = "2.2cm") |>
  column_spec(2, width = "2.5cm") |>
  column_spec(3, width = "5.5cm") |>
  column_spec(4, width = "1.3cm") |>
  column_spec(5, width = "3cm")
```

Three of the four registry first-appearance dates either match the
fuel-station-specific corroborated date exactly (Coomera) or lack any
fuel-station-specific external date for comparison (Casuarina, Lake
Macquarie); these retain the registry date. **Perth Airport is the only
revision:** the ACAPMAg trade-press report explicitly distinguishes the
petrol station (open 19 Feb 2020, bowser price quoted) from the
warehouse (opening one month later on 19 Mar 2020), placing the
fuel-station opening 52 days before the first FuelWatch record.

**Station coordinates and distance-to-Costco.** Distance is computed as
the great-circle distance from each station's coordinates to the nearest
Costco fuel station's coordinates. NSW and WA historical registry files
do not carry lat/lng, so coordinates are joined from one-time pulls of
the live FuelCheck and FuelWatch APIs by (normalized name, postcode);
QLD's historical files include coordinates inline. Stations that fail
the coordinate join are dropped from analysis — they cannot be
classified into a 5 km ring or a donor postcode without a known
location. Per-registry join failure rates are in §2.6.

## Aggregation

**Why monthly.** Three considerations point at month as the right level
of aggregation. First, NSW and QLD are event-based registries (rows are
price-change events) while WA is a daily-snapshot registry; month is
the lowest common frequency at which all three speak the same language
without having to invent 'daily' observations for the event-based feeds.
Second, the research question is about *sustained* competitive response
to Costco entry, which plays out over weeks-to-months rather than days,
so daily resolution would add noise without identification. Third,
monthly means across multiple stations in a ring or postcode smooth
single-quote outliers without requiring any winsorisation step.

**Two-stage aggregation.** Within each station × month, we compute a
single mean unleaded price across all of that station's price
observations in the month. We then take a simple (unweighted) mean
across stations to produce the (ring × month) treated panel and the
(postcode × month) donor panel. The two-stage structure prevents
stations that change prices more often (and therefore appear more times
in the event-based NSW/QLD feeds) from contributing disproportionately
to the cell mean.

**Carry-forward step (NSW and QLD only).** For NSW and QLD, where rows
are price-change events rather than daily snapshots, we construct each
station's daily price series by carrying its last-observed price forward
to subsequent days within the month until the next observed change.
This is the standard approach for converting price-event data to a
daily series and then to monthly means. WA's daily-snapshot data passes
through unchanged.

## Data merging

Three sources are joined into a single analysis panel:

- **State registries** (NSW FuelCheck, QLD FPRS, WA FuelWatch) provide
  the price observations.
- **Live-API station snapshots** (NSW, WA) and inline coordinates (QLD)
  provide station latitude / longitude, joined to historical registry
  rows by (normalized station name, postcode).
- **Hand-collected Costco metadata** (`data/catalogs/usable_costcos.csv`)
  provides each Costco's lat/lng and validated treatment date, joined
  to treated panels by Costco name.

**Expected losses in the merge.** Two sources of leakage. First,
stations that closed, rebranded, or were re-numbered before the April
2026 live-API pull are absent from that snapshot, so historical rows
that mention them fail the coordinate join. The snapshot misses **2,376
of 4,543 NSW historical stations (52.3%)** and **673 of 1,423 WA
historical stations (47.3%)**; QLD is unaffected because its historical
files carry coordinates inline. The loss falls almost entirely on the
donor pool — all four treated Costcos join cleanly — and
`r n_distinct(donor_pool$postcode)` donor postcodes survive the
downstream geographic and quality filters. Second, station-name typos
and mid-panel renames occasionally break the join; the build pipeline
addresses these programmatically where it can (the May 2022 NSW Costco
"Members only" rename, for example, is stripped before matching).

## Filters and transformations: the funnel from raw to analysis-ready

The funnel below traces the exact row and cell counts from the raw
registry pulls through to the four analysis-ready files in
`data/sc_inputs/`. Row counts in the upper block give way to (station ×
month) cells, then (postcode × month) cells, then postcode counts, then
Costco counts as the panel is aggregated and filtered. **Source:**
[`scripts/synthetic_control_input/build_sc_inputs.py`](../scripts/synthetic_control_input/build_sc_inputs.py).
Bolded rows are the milestone outputs.

```{r tbl-funnel}
tibble::tribble(
  ~Step,                                      ~NSW,         ~QLD,         ~WA,         ~Justification,
  "Raw registry pull (rows)",                 "4,401,971",  "4,970,650",  "8,047,153", "Source data",
  "Filter to regular unleaded (rows)",        "1,982,507",  "997,680",    "2,059,755", "Question is about retail unleaded; premium/diesel are different markets",
  "Drop sentinel prices 50–500 ¢/L (rows)",   "1,982,506",  "995,145",    "2,059,755", "Out-of-stock placeholder removal; nearly nonbinding",
  "Join station coordinates (rows)",          "1,247,728",  "995,145",    "1,289,912", "Required for radius assignment; NSW/WA from live-API snapshot",
  "Aggregate to (station × month) cells",     "156,864",    "70,564",     "68,332",    "Carry-forward for NSW/QLD events; passthrough for WA",
  "Aggregate to (postcode × month) cells",    "42,307",     "18,186",     "18,089",    "Donor candidate panel (561 / 245 / 198 postcodes)",
  "Drop postcodes ≤ 20 km from any Costco",   "-231 pcds",  "-51 pcds",   "-80 pcds",  "Donors must be outside any Costco's competitive footprint",
  "Drop postcodes < 3 stations / mo avg",     "-226 pcds",  "-138 pcds",  "-82 pcds",  "Monthly mean too noisy below 3 stations",
  "Drop postcodes < 24 months observed",      "-0 pcds",    "-0 pcds",    "-0 pcds",   "Coverage was good; no postcode failed this gate",
  "Final donor pool (postcodes)",             "104",        "56",         "36",        "196 total × ~90 months = 17,578 (postcode × month) cells",
  "Costco eligibility (≥ 24 pre, ≥ 12 post)", "-4 Costcos", "-2 Costcos", "0",         "Inclusion rule applied to all 10 Australian Costco fuel stations",
  "Final treated panel (Costcos)",            "1",          "1",          "2",         "4 Costcos × 5 km ring × ~80–100 months = 375 (Costco × month) cells"
) |>
  kbl(booktabs = TRUE, align = "lllll",
      caption = "Sample-construction funnel from raw registry rows to the analysis-ready panels in data/sc\\_inputs/.") |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9) |>
  column_spec(1, width = "5cm") |>
  column_spec(2, width = "1.7cm") |>
  column_spec(3, width = "1.5cm") |>
  column_spec(4, width = "1.5cm") |>
  column_spec(5, width = "5.5cm") |>
  row_spec(c(10, 12), bold = TRUE)
```

**Output.** Four CSVs in `data/sc_inputs/`:
`treated_units.csv` (`r comma(nrow(treated_units))` Costco-month rows),
`donor_pool.csv` (`r comma(nrow(donor_pool))` postcode-month rows),
`treated_metadata.csv` (`r nrow(treated_metadata)` rows), and
`donor_metadata.csv` (`r nrow(donor_metadata)` rows). These four files,
uploaded to the Dropbox folder referenced in the setup chunk, form the
complete input to §3's synthetic-control analysis.

\newpage

# Section 3: Synthetic-Control Analysis

## Research design

**Question type.** Causal. The identification strategy is **synthetic
control**, fit separately for each of the four treated Costcos.

**Why synthetic control over difference-in-differences.** With four
treated Costcos and 196 candidate donor postcodes, a treated-versus-control
mean comparison is too thin: a single anomalous Costco (e.g. Casuarina,
with only ~3 contributing stations) would dominate the average. Synthetic
control builds a weighted counterfactual for each treated unit
individually — exactly the design the n = 4 case calls for.

**Source of variation.** Costco fuel-station entry into a local market
isn't driven by what's happening to local fuel prices. Site selection
is driven by Costco warehouse logistics, land availability, and Costco's
national rollout schedule — none of which respond to the local
fuel-price trajectory at the chosen site. The variation we exploit is
the staggered timing of the four openings (Feb 2020, May 2022, Nov 2022,
May 2023) across three states.

**Identifying assumptions.** Each is paired with the figure or check
that defends it:

- *Pre-period fit captures unobserved time-varying confounders.*
  Verified visually in @fig-sc-trajectories (pre-period overlap between
  treated and synthetic), quantified in @tbl-fit-quality (pre-period
  tracking error per Costco), and tested by the 12-month holdout in §4
  Robustness Check 5.
- *Donors stay clean.* No donor postcode is partly affected by a Costco
  that isn't the one we're studying. Defended by the 20 km donor buffer
  (4× the treated radius), pre-committed in §2.2, and explicitly tested
  by §4 Robustness Check 7 (spatial placebo on the 5–20 km donut).
- *No anticipation.* Competitor stations do not preemptively cut prices
  ahead of Costco's announced opening. Tested by §4 Robustness Check 1
  (placebo in time, −12 months).
- *Range coverage.* The treated ring's pre-period price level can be
  expressed as a weighted average of donor postcodes. Verified by the
  donor-weight bars in @fig-donor-weights (no negative weights, no
  extreme single-donor concentration).

**What the weights do.** One synthetic-control fit per treated Costco.
For each Costco the donor pool is restricted to **same-state postcodes
only** so the synthetic absorbs state-specific shocks (state excise
rates, regional refining capacity, intra-state shipping costs) by
construction. Weights are chosen to minimize the average gap between
the treated 5 km ring's monthly mean unleaded price and the weighted
donor mean across the pre-period.

**Method commitment.** Standard synthetic-control via R's `tidysynth`
package — the modern tidyverse-friendly successor to the original
`Synth` implementation. **Predictors:** yearly means of the outcome
series over the pre-period (no auxiliary covariates — we have no
postcode-level demographic panel that would justify them).
**Pre-period objective:** minimize the average gap between treated and
synthetic on monthly mean unleaded price (`optimization_window =
pre_period` in `tidysynth::generate_weights`). **Inference:** estimate
uncertainty by treating each donor postcode as a fake treated unit and
measuring the resulting fake effect, summarized as the post/pre
tracking-error ratio and as a 95% CI on the post-treatment gap. All
fitting is driven by [`_sc_helpers.R::fit_one_costco()`](_sc_helpers.R).

## Fitting the four synthetic controls

```{r sc_fits}
#| cache: true
sc_fits <- treated_metadata |>
  arrange(costco_key) |>
  mutate(
    fit = pmap(
      list(costco_key, treatment_date, state),
      function(name, td, st) {
        fit_one_costco(
          treated_units  = treated_units,
          donor_pool     = donor_pool,
          costco_name    = name,
          treatment_date = td,
          state          = st
        )
      }
    )
  )

trajectories <- sc_fits |>
  transmute(
    costco          = costco_key,
    state,
    treatment_date,
    trajectory      = map2(fit, costco_key, trajectory)
  )

per_costco_pointwise_ci <- sc_fits |>
  transmute(
    costco         = costco_key,
    treatment_date,
    ci             = map(fit, permutation_ci_pointwise)
  )
```

The chunk above fits four `tidysynth` models (one per treated Costco),
each with `generate_placebos = TRUE` so the donor-permutation
distribution is computed in the same pass. Caching is enabled
(`cache: true`) so subsequent knits re-use the fits.

## Synthetic-control trajectories per Costco

Four stacked panels of treated vs synthetic trajectories. Solid colored =
actual treated 5 km ring; dashed black = synthetic counterfactual;
vertical dotted red = validated treatment date (§2.4); shaded gray
band = 95% donor-permutation CI on the synthetic, computed by treating
each donor postcode as a placebo treated unit and taking the empirical
2.5–97.5 percentile of donor-side gaps at each time point. The wider
the band, the larger the noise floor at that month — a treated curve
that falls outside the band is evidence the gap is bigger than what
donor-side variability would produce.

```{r fig-sc-trajectories}
#| fig-cap: "Synthetic-control trajectories per treated Costco, stacked for full-width time comparison. Solid colored = treated 5 km ring (actual); dashed black = synthetic counterfactual; vertical dotted red = validated Costco opening; shaded gray = 95% donor-permutation CI on the synthetic."
#| fig-height: 9
#| fig-width: 7

costco_colors <- c(
  "Coomera"        = "#4C72B0",
  "Casuarina"      = "#DD8452",
  "Perth Airport"  = "#55A868",
  "Lake Macquarie" = "#C44E52"
)

make_traj_panel <- function(costco_name) {
  meta <- treated_metadata |> filter(costco_key == costco_name)
  traj <- trajectories |>
    filter(costco == costco_name) |>
    pull(trajectory) |> pluck(1)
  ci <- per_costco_pointwise_ci |>
    filter(costco == costco_name) |>
    pull(ci) |> pluck(1)

  traj_with_ci <- traj |>
    left_join(ci, by = "time_unit") |>
    mutate(
      synth_lo = synth_y + ci_lo,
      synth_hi = synth_y + ci_hi
    )

  ggplot(traj_with_ci, aes(x = time_unit)) +
    geom_ribbon(aes(ymin = synth_lo, ymax = synth_hi),
                fill = "gray80", alpha = 0.5) +
    geom_line(aes(y = synth_y), linetype = "dashed", color = "black",
              linewidth = 0.5) +
    geom_line(aes(y = real_y), color = costco_colors[[costco_name]],
              linewidth = 0.7) +
    geom_vline(xintercept = as.Date(meta$treatment_date),
               linetype = "dotted", color = "#C44E52", linewidth = 0.6) +
    scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
    labs(x = NULL, y = "Mean unleaded price (¢/L)",
         title = glue("{costco_name} ({meta$state})"),
         subtitle = glue("opens {format(as.Date(meta$treatment_date), '%b %Y')}")) +
    theme_minimal(base_size = 9) +
    theme(plot.title    = element_text(size = 10, face = "bold"),
          plot.subtitle = element_text(size = 8,  color = "gray40"))
}

traj_panels <- map(costco_order, make_traj_panel)
wrap_plots(traj_panels, ncol = 1)
```

A reader who sees only @fig-sc-trajectories should be able to read
off the project's conclusion: did competitor prices fall after Costco
entry, and by how much.

## Donor weights per treated Costco

For each Costco we report the postcodes whose synthetic-control weight
exceeds 0.01, sorted in descending order. A counterfactual leaning on
one or two donors is fragile; one drawn from a balanced mix of
geographically sensible donors is credible. Geographic sanity-check:
Coomera should lean on Gold Coast / Brisbane postcodes, Lake Macquarie
on Newcastle / Central Coast, Perth Airport and Casuarina on Perth
metro postcodes.

```{r fig-donor-weights}
#| fig-cap: "Donor-postcode weights per treated Costco (weights > 0.01, sorted descending). A balanced mix across 5-7+ donors with no single donor exceeding ~0.5 is the credibility signal; concentration on 1-2 donors would flag a fragile counterfactual."
#| fig-height: 7
#| fig-width: 8

weights_long <- sc_fits |>
  transmute(
    costco  = costco_key,
    weights = map(fit, donor_weights_top, threshold = 0.01)
  ) |>
  unnest(weights) |>
  mutate(postcode = sub("^pc_", "", unit))

make_weights_panel <- function(costco_name) {
  d <- weights_long |>
    filter(costco == costco_name) |>
    arrange(desc(weight)) |>
    mutate(label = factor(postcode, levels = rev(postcode)))

  if (nrow(d) == 0) {
    return(
      ggplot() + theme_void() +
        labs(title = glue("{costco_name}: no donors with weight > 0.01"))
    )
  }

  ggplot(d, aes(x = weight, y = label)) +
    geom_col(fill = costco_colors[[costco_name]], alpha = 0.85) +
    geom_text(aes(label = sprintf("%.2f", weight)),
              hjust = -0.2, size = 2.8) +
    scale_x_continuous(expand = expansion(mult = c(0, 0.15)),
                       limits = c(0, NA)) +
    labs(x = "Synthetic-control weight", y = "Postcode",
         title = glue("{costco_name}")) +
    theme_minimal(base_size = 9) +
    theme(plot.title = element_text(size = 10, face = "bold"))
}

weight_panels <- map(costco_order, make_weights_panel)
wrap_plots(weight_panels, ncol = 2)
```

## Aggregate event study

The per-Costco picture in @fig-sc-trajectories obscures the aggregate
dynamic. @fig-event-study turns "Costco lowered prices on average" into
"Costco lowered prices by *X* ¢/L within *Y* months and the effect
persisted for *Z*"
— much closer to the answer a regulator can act on. The window is
−24 to +24 relative months, the largest range over which all four
Costcos contribute (limited by Perth Airport's 26 pre-months and
Coomera's 31 post-months).

```{r fig-event-study}
#| fig-cap: "Aggregate event study. X-axis: months relative to Costco opening. Y-axis: mean of (actual - synthetic) gap across the four treated Costcos. Shaded gray: 95% CI from cross-Costco standard error. Vertical dotted red: treatment (month 0)."
#| fig-height: 4.5
#| fig-width: 7

rel_month_gaps <- trajectories |>
  mutate(
    rel = map2(trajectory, treatment_date, relative_month_gaps)
  ) |>
  select(costco, state, rel) |>
  unnest(rel) |>
  filter(rel_month >= -24, rel_month <= 24)

agg_event <- rel_month_gaps |>
  group_by(rel_month) |>
  summarize(
    n        = sum(!is.na(gap)),
    mean_gap = mean(gap, na.rm = TRUE),
    se       = sd(gap, na.rm = TRUE) / sqrt(sum(!is.na(gap))),
    .groups  = "drop"
  ) |>
  mutate(
    ci_lo = mean_gap - 1.96 * se,
    ci_hi = mean_gap + 1.96 * se
  )

ggplot(agg_event, aes(x = rel_month, y = mean_gap)) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi),
              fill = "gray80", alpha = 0.6) +
  geom_hline(yintercept = 0, color = "gray40", linewidth = 0.3) +
  geom_vline(xintercept = 0, linetype = "dotted",
             color = "#C44E52", linewidth = 0.5) +
  geom_line(linewidth = 0.7, color = "#4C72B0") +
  geom_point(size = 1.2, color = "#4C72B0") +
  labs(
    x        = "Months relative to Costco opening",
    y        = "Mean gap (actual - synthetic), ¢/L",
    title    = "Aggregate event study: mean gap by month-since-opening",
    subtitle = "Averaged across the 4 treated Costcos; 95% cross-Costco CI"
  ) +
  theme_minimal(base_size = 11)
```

## Per-Costco effect-size summary

```{r tbl-effect-summary}
table_1 <- pmap_dfr(
  list(sc_fits$fit, sc_fits$costco_key, sc_fits$state,
       sc_fits$treatment_date, sc_fits$n_post_months),
  effect_summary
)

table_1 |>
  mutate(ci_str = sprintf("[%.2f, %.2f]", ci_lo, ci_hi)) |>
  transmute(
    Costco             = costco,
    State              = state,
    `Pre mean (¢/L)`   = round(pre_mean_cents, 2),
    `Post gap (¢/L)`   = round(mean_post_gap_cents, 2),
    `\\% red.`         = round(pct_reduction, 2),
    `MSPE ratio`       = round(post_pre_ratio, 2),
    `95\\% CI`         = ci_str,
    `Post mo.`         = post_months
  ) |>
  kbl(booktabs = TRUE, align = "llrrrrrr", escape = FALSE,
      caption = "Per-Costco effect-size summary. Mean post-treatment gap = mean of (actual - synthetic) across the post-treatment window. Post/pre MSPE ratio compares post-period mean-squared prediction error to pre-period; values of 4-5+ are the standard threshold for evidence of a real effect.") |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9) |>
  column_spec(1, width = "2.4cm") |>
  column_spec(2, width = "0.8cm") |>
  column_spec(3, width = "1.7cm") |>
  column_spec(4, width = "1.7cm") |>
  column_spec(5, width = "1.2cm") |>
  column_spec(6, width = "1.6cm") |>
  column_spec(7, width = "2.5cm") |>
  column_spec(8, width = "1.2cm")
```

The **post/pre tracking-error ratio** compares how far the synthetic
strays from actual *after* Costco entry to how closely it tracked
*before*. A ratio of 4–5 or higher is the standard threshold for
evidence of a real effect — the post-period gap is several times larger
than the pre-period noise. The **95% CI** is the empirical 2.5–97.5
percentile of post-treatment mean gaps across the donor-permutation
distribution; a CI that does not include zero is evidence against the
null of "no effect."

## Pre-period fit quality

```{r tbl-fit-quality}
table_2 <- pmap_dfr(
  list(sc_fits$fit, sc_fits$costco_key, sc_fits$treatment_date,
       sc_fits$n_pre_months),
  fit_quality_summary
)

table_2 |>
  transmute(
    Costco                              = costco,
    `Pre months`                        = pre_months,
    `Pre tracking error (¢/L)`     = round(pre_rmspe_cents, 3),
    `Tracking error as \\% of pre mean` = round(pre_rmspe_pct, 2),
    `Top-3 donor weight share`          = round(top3_donor_share, 2)
  ) |>
  kbl(booktabs = TRUE, align = "lrrrr", escape = FALSE,
      caption = "Pre-period fit quality per treated Costco. Lower tracking error means a tighter pre-period match; a tracking error below 1\\% of the pre-period mean is typically considered a credible fit. Top-3 donor weight share reflects how concentrated the synthetic is on its strongest donors.") |>
  kable_styling(latex_options = c("striped", "hold_position"),
                font_size = 9) |>
  column_spec(3, width = "3cm") |>
  column_spec(4, width = "3.5cm") |>
  column_spec(5, width = "3cm")
```

Casuarina's pre-period tracking error is expected to be visibly larger
than the others because the 5 km Casuarina ring contains only ~3
contributing stations on average (§1's documented limitation).
Surfacing this transparently here is part of presenting the limitation
rather than hiding it.

\newpage

# Section 4: Robustness Checks

> **This section is stubbed for the first knit.** Once §3 lands, we'll add
> the seven pre-committed robustness checks from
> `deliverables/plan_of_attack_section_3.pdf` §(c):
>
> 1. Placebo in time (−12 months).
> 2. Alternative radii (3 km / 15 km and 8 km / 30 km).
> 3. Casuarina at 10 km treated radius.
> 4. Perth Airport with Mar–Dec 2020 dropped.
> 5. LOOCV donor selection + 12-month holdout.
> 6. 95% CIs via donor-permutation inference (also drives @fig-sc-trajectories bands).
> 7. Spatial placebo: 5–20 km donut as faux donors.
>
> Checks 2, 3, and 7 require alternate-radius SC inputs, which we will
> build by parameterizing `scripts/synthetic_control_input/build_sc_inputs.py`
> and re-running it at the new geometries; the resulting CSVs join the
> Dropbox folder.

\newpage

# Section 5: Recommendation

> **This section is stubbed for the first knit.** Once the headline effect
> has a number, we'll translate it into a concrete ACCC recommendation and
> a back-of-envelope estimate of aggregate consumer savings per Costco
> entry (effect size × ring volume × annual months).

\newpage

# Section 6: Limitations

> **This section is stubbed for the first knit.** Planned coverage:
>
> - **n = 4 treated Costcos.** Headline averages cross four units; the
>   per-Costco point estimates carry more weight than the pooled mean.
> - **Casuarina sparsity.** ~3 stations contribute to the 5 km ring on
>   average; the per-Costco gap is noisier than for the others. Robustness
>   Check 3 probes whether widening to 10 km changes the picture.
> - **Perth Airport / COVID confound.** Treatment falls four weeks before
>   the demand collapse; Robustness Check 4 isolates the pandemic period.
> - **ACT excluded.** Canberra Airport Costco opened in Nov 2023 but the
>   ACT only joined NSW FuelCheck after the opening, leaving no pre-period
>   to fit on. Six other Costcos are excluded for analogous data-coverage
>   reasons (plan-of-attack §1.4 / §2(b)).
> - **Same-state donor pools only.** Cross-state donors would require
>   absorbing state-level shocks (excise rates, refining capacity,
>   shipping costs) which the within-state restriction handles by
>   construction. The cost is a smaller donor pool, especially for QLD and WA.

\newpage

# Appendix: how this analysis is reproduced

This `.qmd` reads its five input CSVs from a public Dropbox folder, fits
synthetic-control models with `tidysynth`, and renders to PDF in a single
`quarto render` call. The full reproduction procedure — including required
R packages, the Dropbox folder URL, and the location of upstream pipeline
scripts — is documented in `analysis/README.md`.
